---
title: "Leach BLCN Report Data And Mapping Compilation"
author: Andrew Leach
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      code_folding: hide
      includes:
      after_body: 
      theme: lumen
always_allow_html: yes
editor: source

---
```{r, echo=FALSE,include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,cache = T,fig.align='center',dpi=220,fig.width=16,fig.height=9)
options(scipen = 999)
```



```{r,cache=F, include=F}
library(janitor)
library(scales)
library(tidyverse)
library(readxl)
library(openxlsx)
library(viridis)
library(ggthemes)
library(kableExtra)
library(sf)
library(cowplot)
library(ggrepel)
library(rvest)

caf_blue<-"#003080"

work_theme<-function(){
  theme_tufte(18)+
     theme(
           plot.subtitle = element_text(color="grey10",size=rel(1.25)),
           plot.title = element_text(face="bold"),
           plot.caption = element_text(color="grey50",size=rel(1)),
           legend.title = element_text(color="grey10",size=rel(1)),
           legend.text = element_text(color="grey10",size=rel(1)),
           strip.text = element_text(size=rel(1.2)),
           axis.title = element_text(size=rel(1.2)),
           axis.text = element_text(size=rel(1.2)),
           axis.text.x = element_text(size=rel(.8)),
           axis.ticks = element_blank(),
           panel.spacing = unit(.75,"lines"),
           legend.position = "bottom",
           plot.margin = margin(t = .5, r = .5, b = .5, l = .5,unit= "cm"),
           #axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)
           NULL
           )+
  NULL
}

colors_tableau10<-function () 
{
    return(c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", 
        "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17BECF"))
}

```


# Territory Shape Files

```{r clawr}
library(httr)
#set crs
crs_val<-3347
url <- "https://geospatial.alberta.ca/titan/rest/services/boundary/military_base_and_training_ground/MapServer/0/query"
# Set up the query parameters
params <- list(
  where = "1=1",
  outFields = "*",
  f = "geojson"
)

# Make the GET request
response <- GET(url, query = params)
# Check if the request was successful
if (status_code(response) == 200) {
  # Parse the GeoJSON content
  geojson_content <- content(response, as = "text", encoding = "UTF-8")
  clawr_data <- st_read(geojson_content,quiet = TRUE)%>%
  st_transform(crs_val)
} else {
  stop("Failed to retrieve data from the REST service.")
}


```


```{r reserves and cities,fig.align='center',fig.width=16, fig.height=8,dpi=220}

#gd_pts <- st_read("../blcn/Pipelines_SHP/Pipelines_GCS_NAD83.shp")%>%clean_names()

res_pts <- st_read("../blcn/ILC/Indigenous_Lands_of_Canada.shp",quiet=TRUE)%>%clean_names()%>%
  st_transform(crs_val)%>%
  filter(grepl("BEAVER LAKE 131",admin_area_n))


cities<- st_read("../blcn/Municipal_Boundaries_SHP_Geographic/CITY.shp",quiet=TRUE)%>%
   filter(grepl("COLD",GEONAME))%>%
  bind_rows(st_read("../blcn/Municipal_Boundaries_SHP_Geographic/TOWN.shp",quiet=TRUE)%>%
              filter(grepl("ATHABASCA",GEONAME)))%>%
   st_transform(crs_val)


l_ath_bd<- st_read("../blcn/LARP/LUF Integrated Regional Plan Boundaries.shp",quiet=TRUE)%>%
  filter(LUF_CODE=="02")%>%
  st_transform(crs_val)

ats<- st_read("../blcn/ATS_Polygons_SHP_Geographic/V4-1_TWP.shp",quiet=TRUE)%>%
  filter(M==4)%>%
  group_by(TWP,RGE)%>%
  summarize(geometry = st_union(geometry), .groups = "drop")%>%
  st_transform(crs_val)


larp_pop<- st_read("../blcn/LARP/LA_RAC_Recommended_Population_Centers.shp",quiet=TRUE)%>%
  st_transform(crs_val)

blcn_terr <-st_read("../blcn/blcn_poly/Beaver Lake Cree Traditional Territory.shp",quiet=TRUE)%>%
  st_transform(crs_val)


poly <- st_transform(blcn_terr, crs = 4326)
dir.create("shapefile_output", showWarnings = FALSE)
st_write(poly, "shapefile_output/blcn_polygon.shp", delete_layer = TRUE)
zip::zipr("blcn_polygon.zip", files = list.files("shapefile_output", full.names = TRUE))

# Buffer in meters (20 km = 20,000 meters)
blcn_buffer <- st_buffer(blcn_terr, dist = 20000)


blcn_ats<-st_filter(ats, blcn_terr, .predicate = st_intersects)

twp_range<-range(blcn_ats$TWP)
rge_range<-range(blcn_ats$RGE)

blcn_ats<-ats %>%
  filter(RGE%in%c(1:23))%>%
  filter(TWP%in%c(59:82))
  

blcn_ats_strict<-st_intersection(blcn_terr, ats)


# Buffer in meters (500 km = 500,000 meters)
blcn_buffer_500 <- st_buffer(blcn_terr, dist = 500000)
blcn_buffer_100 <- st_buffer(blcn_terr, dist = 100000)

# Buffer in meters (50 km = 50,000 meters)
blcn_buffer_50 <- st_buffer(blcn_terr, dist = 50000)

ats_first_range <- blcn_ats %>%
  filter(RGE == 1)

# Filter ATS to townships >= 67
ats_filtered <- blcn_ats %>% filter(TWP >= 67)

ats_union <- st_union(ats_filtered)

# Intersect BLCN territory with the ATS boundary (clip)
blcn_67p <- st_intersection(blcn_terr, ats_union)
# Plot and label only those townships
ggplot(blcn_terr) +
  #geom_sf(data=ngtl_blcs)+
  geom_sf(data = blcn_ats_strict, fill = NA, color = "grey80") +  # full grid for context
  #geom_sf(data = ats_first_range%>%filter(TWP>=67), fill = NA, color = "black") +  # highlight R1
  geom_sf(aes(colour="Beaver Lake Cree Nation Core Territory (trimmed to TWP 67)"),fill=NA,linewidth=0.85,lty="11") +
  scale_color_manual("",values=c("black"))+
  #  geom_sf_text(
  #  data = ats_first_range,
  #  aes(label = TWP),
  #  size = 3,
  #  color = "black"
  # ) +
  theme_minimal()+
theme(panel.grid = element_blank(),
        legend.position = "none",
        axis.text = element_blank(),
      axis.title = element_blank())

#ngtl_blcs<-st_filter(ngtl, blcn_buffer_50, .predicate = st_intersects)
ggplot(blcn_terr)+
  #geom_sf(data = blcn_ats,size=2,fill=NA)+
  geom_sf(aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=0.85,colour="black",lty="11") +
  theme_minimal()+
  theme(panel.grid = element_blank(),
        legend.position = "none",
        axis.text = element_blank())+
  
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,8,11)),size=2)+
  geom_sf(data = cities,size=2)+
  
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(3,6,  8, 11)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 2
  )
ggsave("test.png",bg="transparent",dpi=300)


```


```{r,fig.width=12,fig.height=10,dpi=300}
# ArcGIS FeatureServer query endpoint
url <- "https://services2.arcgis.com/jQV6VMr2Loovu7GU/arcgis/rest/services/Oil_Sands_Area/FeatureServer/0/query"

# Query parameters
params <- list(
  where = "1=1",        # select all
  outFields = "*",      # all attributes
  f = "geojson"         # output format
)

# Download the GeoJSON directly from ArcGIS REST API
res <- GET(url, query = params)

# Parse and convert to sf
oil_sands_area <- st_read(content(res, as = "text"), quiet = TRUE) %>%st_transform(crs_val)

#save(oil_sands_area,file="oil_sands_area.rdata")

library(rnaturalearth)
library(rnaturalearthdata)

# get province boundary
canada <- ne_states(country = "canada", returnclass = "sf")
alberta <- canada %>% filter(name_en == "Alberta")

```

```{r,fig.width=8,fig.height=6,fig.dpi=300}

target_crs <- 26912


ggplot(oil_sands_area%>%
         #mutate(Area_Name=ifelse(grepl("Peace",Area_Name),"Peace River",Area_Name))%>%
         st_transform(target_crs)
       )+
  geom_sf(data = alberta%>%st_transform(target_crs),
          fill = "grey95", colour = "grey70", linewidth = 0.5) +
  geom_sf(aes(fill=Area_Name),linewidth=0.25,alpha=0.6)+
  geom_sf(data=blcn_terr%>%st_transform(target_crs),aes(colour="Beaver Lake Cree Nation\nCore Territory"),fill=NA,linewidth=1.5,lty="solid") +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data%>%st_transform(target_crs),aes(colour="Cold Lake Air Weapons\nRange (CLAWR)"),
          fill=NA,linewidth=1.5,lty="solid")+
  scale_color_manual("",values=c("black",caf_blue,"grey60"))+
  scale_fill_viridis("Oil Sands Area",discrete = T,option = "cividis",begin = 0.2,end = 0.8)+
  guides(fill = guide_legend(order = 2))+
  theme_void()+
  theme(
    legend.position = "right",
    legend.box = "vertical",
    text = element_text(size = 12,face = "bold"),
    legend.title = element_text(size = 10,face = "bold"),
    plot.margin = margin(t = 5, r = 5, b = 5, l = 5)  # right margin expanded (r = 50)
    )+
    labs(x="",y="",title="",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       #caption="Data via Alberta Energy Regulator ST44, map by Andrew Leach",
       NULL
       )

ggsave("os_regions.png",width = 7,height = 7,dpi=300,bg="white")

```





```{r,fig.width=12,fig.height=10,dpi=300}
#oil sands royalty projects from AB Map

url <- "https://gis.energy.gov.ab.ca/arcgis/rest/services/Geoview/Royalty_Ext_PROD/MapServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

os_project_map_base <- st_read(url)%>%st_transform(crs_val)
#save(os_project_map,file="os_project_map_all.rdata")
```



```{r proj_map,fig.align='center',fig.width=16, fig.height=8,dpi=220}

if(!file.exists("ab_regimes.zip")){
download.file("https://www.aer.ca/data/shapefiles/Scheme_Approval_SHP.zip",destfile = "ab_regimes.zip",mode="wb")
unzip("ab_regimes.zip")
}


in_situ_shp <- st_read("Shapefiles/In Situ Oil Sands Scheme Approvals/In Situ Oil Sands Scheme Approval_GCS_NAD83.shp",quiet=TRUE)%>%clean_names()%>%
  st_transform(crs=crs_val)
  

#AER ST 102 Facilities List
if(!file.exists("ST102.zip")){
download.file("https://www.aer.ca/data/codes/ST102-SHP.zip","ST102.zip",mode="wb")
unzip("ST102.zip")}

ST102<-st_read("ST102_Facility_GCS_NAD83.shp",quiet=TRUE)%>%clean_names()%>%
  st_transform(crs_val)


#select for schemes in BLCN territory
terr_fac <- st_filter(ST102, blcn_terr, .predicate = st_intersects)%>%distinct()%>%
    #mutate(scheme_num=str_extract(scheme_no, "^\\d+"),
    #       sche_name = replace(sche_name, is.na(sche_name), ""))%>%
  I()


#EPA Oil sands Projects
  
if(!file.exists("epa_oil_sands.xlsx"))
   download.file("https://open.alberta.ca/dataset/9bc110e1-37ec-4264-8c7f-cb678e288e44/resource/5d54e2bd-2dcb-4eb6-9b8b-ffe23fb01305/download/epa-oil-sands-operating-capacity-report-2023-2024.xlsx",dest="epa_oil_sands.xlsx",mode="wb")


#download.file("https://www.aer.ca/prd/documents/sts/ST44_2024-Q4.xlsx",dest="aer_os_schemes.xlsx",mode="wb")


#AER ST44 Oil sands Schemes Excel
  
if(!file.exists("aer_os_schemes.xlsx"))
  download.file("https://www.aer.ca/prd/documents/sts/ST44_2024-Q4.xlsx",dest="aer_os_schemes.xlsx",mode="wb")

aer_active_schemes<-read_excel("aer_os_schemes.xlsx",sheet="ActiveInSitu",skip=3)%>%clean_names()%>%
  rename(sche_name=scheme_name)%>%
  mutate(active=TRUE)

aer_inactive_schemes<-read_excel("aer_os_schemes.xlsx",sheet="Inactive",skip=3)%>%clean_names()%>%
  rename(sche_name=scheme_name)%>%
  mutate(active=FALSE)

aer_schemes<-bind_rows(aer_active_schemes,aer_inactive_schemes)%>%
  mutate(scheme_num=str_extract(approval_number, "^\\d+"))%>%
  distinct()


```



```{r insitu,fig.width=14,fig.height=7,dpi=300}
#select for schemes in BLCN territory
in_situ <- st_filter(in_situ_shp, blcn_terr, .predicate = st_intersects)%>%distinct()%>%
    mutate(scheme_num=str_extract(scheme_no, "^\\d+"),
           sche_name = replace(sche_name, is.na(sche_name), ""))


#exp<-in_situ %>% select(scheme_no,scheme_num,scheme_typ,approval_number=app_no,app_holder,area=geo_field)
#save(exp,file="../st53/blcn_app.rdata")
load(file="../st53/blcn_app.rdata")

load("st53timing.rdata") #loads BLCN_timing


in_situ<-in_situ%>%
    left_join(aer_schemes%>%
        mutate(sche_name = replace(sche_name, is.na(sche_name), "")),
        by=c("scheme_num","sche_name"))%>%
    mutate(status=ifelse(is.na(status),"",status))%>%
      left_join(BLCN_timing%>%rename(first_st53=first_date))%>%
    mutate(
        issue_bin = cut(
          year(issue_date),
          breaks = c(-Inf, 1999, 2010, 2020, Inf),
          labels = c("Pre-2000", "2000-2010", "2010-2020", "2020-present"),
          right = TRUE
        ),
        st53_bin = cut(
          year(first_st53),
          breaks = c(-Inf, 2000, 2010, 2020, Inf),
          labels = c("Producing in 2000", "2001-2010", "2011-2020", "2021-present"),
          right = TRUE
        )
      )

print(NROW(in_situ))
print(NROW(unique(in_situ$scheme_num)))


table(in_situ$status)

#test<-
  in_situ %>% filter(status!="")%>%distinct()%>%
  pull(sch_sub_ty) %>%
  table()

in_situ<-in_situ%>%distinct()%>%
   filter(status!="Superseded")%>%
   filter(status!="Expired")%>%
   filter(status!="Terminated Not Started")


# Frequency table of both bins
in_situ %>% pull(st53_bin)%>% tibble()%>%
  rename(st53_bin=1)%>%
  count(st53_bin) %>%
  arrange(st53_bin)

issue_bin_levels <- c("Pre-2000", "2000-2010", "2010-2020", "2020-present")
bin_levels <- c("Pre-2000", "Pre-2010", "Pre-2020", "Through 2024")
in_situ_2000<-in_situ %>% filter(year(issue_date)<=2000)%>%
  mutate(year="2000")
in_situ_2010<-in_situ %>% filter(year(issue_date)<=2010)%>%
  mutate(year="2010")
in_situ_2020<-in_situ %>% filter(year(issue_date)<=2020)%>%
  mutate(year="2020")
in_situ_present<-in_situ %>% 
  mutate(year="End 2024")

in_situ_prog<-bind_rows(in_situ_2000,in_situ_2010,in_situ_2020,in_situ_present)%>%
  mutate(year=as_factor(year),
         issue_bin=factor(issue_bin,levels = issue_bin_levels)
         )



ggplot(in_situ_prog)+
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1,lty="11") +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  geom_sf(aes(colour="Oil Sands Scheme Approval Boundaries",fill=issue_bin),linewidth=0.05)+
  facet_wrap(~year,nrow=1)+
  scale_color_manual("",values=c("black","khaki","grey60"))+
  scale_fill_viridis("Approval Issue Date",discrete = T,option = "cividis",begin = 0.2,end = 0.8)+
    geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,15)),size=1)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(3,6,7,  8, 15)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    #nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 2,
    #min.segment.length = 0
  )+
 tombe_theme()+
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    legend.title = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  facet_wrap(~year,nrow = 1)+
  labs(x="",y="",title="Oil Sands Scheme Approval Boundaries Within BLCN Territory Over Time",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       #caption="Data via Alberta Energy Regulator ST44, map by Andrew Leach",
       NULL
       )
 

st53_bin_levels <- c("Producing in 2000", "2001-2010", "2011-2020", "2021-present","No production in ST-53")
in_situ_2000<-in_situ %>% filter(year(first_st53)<2001)%>%
  mutate(year="2000")
in_situ_2010<-in_situ %>% filter(year(first_st53)<=2010)%>%
  mutate(year="2010")
in_situ_2020<-in_situ %>% filter(year(first_st53)<=2020)%>%
  mutate(year="2020")
in_situ_present<-in_situ %>% 
  mutate(year="End 2024")

st53_prog<-bind_rows(in_situ_2000,in_situ_2010,in_situ_2020,in_situ_present)%>%
  mutate(year=as_factor(year),
        st53_bin=fct_explicit_na(st53_bin, na_level = "No production in ST-53"),
         first_oil=factor(st53_bin,levels = st53_bin_levels)
         )



ggplot(st53_prog)+
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1,lty="11") +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  geom_sf(aes(colour="Oil Sands Scheme Approval Boundaries",fill=st53_bin),linewidth=0.05)+
  facet_wrap(~year,nrow=1)+
  scale_color_manual("",values=c("black","khaki","grey60"))+
  scale_fill_viridis("First Recorded Production",discrete = T,option = "cividis",begin = 0.2,end = 0.8)+
    geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,15)),size=1)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(3,6,7,  8, 15)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    #nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 2,
    #min.segment.length = 0
  )+
 tombe_theme()+
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    legend.title = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  facet_wrap(~year,nrow = 1)+
  labs(x="",y="",
       #title="Oil Sands Production Commencement by Scheme Within BLCN Territory Over Time",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       #caption="Data via Alberta Energy Regulator ST44, map by Andrew Leach",
       NULL
       )
 


#combine for final map
in_situ<-in_situ%>%
  group_by(sche_name,app_holder) %>%
  summarise(
    first_st53=min(first_st53),
    issue_date=min(issue_date),
    geometry = st_union(geometry),  # combine geometries
    sche_name = first(sche_name),   # keep one name per group
    status = list(status),         # same for other fields
    app_holder = first(app_holder),         # same for other fields
    operator = first(operator),         # same for other fields
    sch_sub_ty = first(sch_sub_ty),         # same for other fields
    .groups = "drop"
  )%>%
  mutate(
    issue_bin = cut(
      year(issue_date),
      breaks = c(-Inf, 1999, 2005, 2010, 2015, 2020, Inf),
      labels = c("pre-2000", "2000-2005", "2005-2010", "2010-2015", "2015-2020", "2020-present"),
      right = TRUE
    ),
    st53_bin = cut(
      year(first_st53),
      breaks = c(-Inf, 2000, 2005, 2010, 2015, 2020, Inf),
      labels = c("Producing in 2000", "2001-2005", "2005-2010", "2010-2015", "2015-2020", "2020-present"),
      right = TRUE
    )
  )


print(NROW(in_situ))
print(NROW(unique(in_situ$scheme_num)))


#test<-
  in_situ%>%filter(status=="")%>%pull(app_holder)%>%unique()%>%tibble()%>%write_csv("firms.csv")

in_situ%>%pull(operator)%>%unique()%>%tibble()%>%
  rename("firm"=1)%>%
  mutate(firm=gsub("Limited","Ltd.",firm),
         firm=gsub("Corporation","Corp.",firm))%>%
  arrange(firm)%>%
  write_csv("op_firms.csv")
```




```{r,fig.width=14,fig.height=9,dpi=300}

terr_schemes <- st_filter(in_situ_shp, blcn_terr, .predicate = st_intersects)%>%
    mutate(scheme_num=str_extract(scheme_no, "^\\d+"),
           sche_name = replace(sche_name, is.na(sche_name), ""))

#os_project_map<-st_filter(os_project_map_base, blcn_buffer_50, .predicate = st_intersects)
os_project_map<-st_filter(os_project_map_base, terr_schemes, .predicate = st_intersects)

#save(os_project_map,file="os_project_map.rdata")
#load(file="os_project_map.rdata")

# 1. Intersect with `terr`
intersects <- lengths(st_intersects(terr_schemes,blcn_terr)) > 0

# 2. Not completely within `terr`
not_within <- lengths(st_within(terr_schemes, blcn_terr)) == 0

# 3. Combine: Features that straddle the boundary
straddling_features <- terr_schemes[intersects & not_within, ]


ggplot(os_project_map%>%select(ProjectName,geometry)%>%distinct())+
  geom_sf(data=in_situ,aes(fill="Other ST44 Regime Boundaries"),linewidth=0.15,colour="black") +
  geom_sf(data=straddling_features,aes(fill=paste("AER Scheme Number",scheme_num)),linewidth=0.15,colour="black") +
  
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1,lty="11") +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  geom_sf(aes(colour="Oil Sands Royalty Project Locations"),linewidth=0.5) +
  
  
  geom_label_repel(
    aes(label = ProjectName,geometry=geometry),size=2.5, color="black",
    stat = "sf_coordinates",
    max.overlaps = Inf,
    min.segment.length = 0,
    segment.size = 0.75,         # thicker lines
    segment.colour="black",
    segment.linetype="11",
    force = 10,                     # increase repelling force between labels
    nudge_x = 10,
    nudge_y = 20,
  force_pull = 0.5,              # reduce pull toward anchor
  box.padding = 0.7,             # more space around labels
  )+
    #geom_sf(data=res_pts,aes(colour="Beaver Lake #131 Reserve"),fill="blue",linewidth=1.5) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(2,3,6)),size=2)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(2,3,6)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 2
  )+
  tombe_theme()+
  scale_color_manual("",values=c(colors_colorblind_10()[1:2],"black"))+
  scale_fill_manual("",values=c(colors_tableau10(),colors_tableau10_light()[1:3],"grey90"))+
  guides(colour=guide_legend(ncol = 1,order = 2),fill=guide_legend(ncol = 4,order = 1))+
  theme(
    legend.position = "bottom",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 10),
    legend.title = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  #facet_wrap(~year,nrow = 1)+
  labs(x="",y="",
       #title="Oil Sands Royalty Projects Within Oil Sands Regimes Intersecting BLCN Territory",
       #caption="Oil Sands Royalty Project Locations via Alberta Energy (https://gis.energy.gov.ab.ca/geoview/Royalty), map by Andrew Leach"
       )


```



```{r,fig.width=13,fig.height=8,dpi=300}

straddle_projects<-st_filter(os_project_map_base, straddling_features, .predicate = st_intersects)%>%
  mutate(GrossRevenue=as.numeric(gsub(",","",GrossRevenue)),
         VolOfCleanCrudeBitumen=as.numeric(gsub(",","",VolOfCleanCrudeBitumen)))




ggplot(straddling_features)+
  geom_sf(aes(fill=paste("AER ST44 Scheme Number",scheme_no)),linewidth=0.15,colour="black") +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1,lty="11") +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  geom_sf(data=straddle_projects,aes(colour="Oil Sands Royalty Project Locations"),linewidth=0.5) +
  
  
  
  geom_label_repel(data=straddle_projects,
    aes(label = ProjectName,geometry=geometry),size=2.5, color="black",
    stat = "sf_coordinates",
    max.overlaps = Inf,
    min.segment.length = 0,
    segment.size = 0.75,         # thicker lines
    segment.colour="black",
    segment.linetype="11",
    force = 10,                     # increase repelling force between labels
    nudge_x = 10,
    nudge_y = 20,
  force_pull = 0.5,              # reduce pull toward anchor
  box.padding = 0.7,             # more space around labels
  )+
    #geom_sf(data=res_pts,aes(colour="Beaver Lake #131 Reserve"),fill="blue",linewidth=1.5) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(2,3,6)),size=2)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(2,3,6)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 2
  )+
  tombe_theme()+
  #scale_color_manual("",values=c(colors_colorblind_10()[1:2],"black"))+
  #scale_fill_manual("",values=colors_colorblind_10()[3:10])+
  guides(colour=guide_legend(nrow = 2,order = 2),fill=guide_legend(nrow = 2,order = 1))+
  theme(
    legend.position = "bottom",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 10),
    legend.title = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  #facet_wrap(~year,nrow = 1)+
  labs(x="",y="",
       #title="Oil Sands Royalty Projects Within Oil Sands Regimes Intersecting BLCN Territory",
       #caption="Oil Sands Royalty Project Locations via Alberta Energy (https://gis.energy.gov.ab.ca/geoview/Royalty), map by Andrew Leach"
       )


```




```{r}

company_levels <- c(
  "Devon","Harvest", "IPC", "Cenovus", "MEG", "Imperial",
  "Athabasca", "CNRL", "Grizzly", "Strathcona",
  "ConocoPhillips",                      # main entries
  "Other Primary and EOR Regimes"       # must come last
)



in_situ<-in_situ%>%
  mutate(sche_name=ifelse(is.na(sche_name),"",sche_name))%>%
  mutate(sche_name=ifelse(sche_name=="Christina Lake (MEG)","Christina Lake",sche_name))%>%
  mutate(app_holder=case_when(
     app_holder=="Harvest Operations Corp." ~ "Harvest",
     app_holder=="Ipc Canada Ltd." ~"IPC",                      
     app_holder=="Cenovus Energy Inc."~"Cenovus",
     app_holder=="MEG Energy Corp."~"MEG",
     app_holder=="Imperial Oil Resources Limited"~"Imperial",
     app_holder=="Athabasca Oil Corporation"~"Athabasca",
     app_holder=="Canadian Natural Resources Limited"~"CNRL",
     app_holder=="Grizzly Oil Sands ULC"~"Grizzly",
     app_holder=="Devon Canada Corporation"~"Devon",
     app_holder=="Strathcona Resources Ltd."~"Strathcona",
     app_holder=="ConocoPhillips Canada Resources Corp."~"ConocoPhillips",
     TRUE~app_holder
  ))%>%
  mutate(operator=case_when(
     operator=="Harvest Operations Corp." ~ "Harvest",
     operator=="Ipc Canada Ltd." ~"IPC",                      
     operator=="Cenovus Energy Inc."~"Cenovus",
     operator=="MEG Energy Corp."~"MEG",
     operator=="Imperial Oil Resources Limited"~"Imperial",
     operator=="Athabasca Oil Corporation"~"Athabasca",
     operator=="Canadian Natural Resources Limited"~"CNRL",
     operator=="Grizzly Oil Sands ULC"~"Grizzly",
     operator=="Strathcona Resources Ltd."~"Strathcona",
     operator=="ConocoPhillips Canada Resources Corp."~"ConocoPhillips",
     TRUE~operator
  ))%>%
  mutate(app_holder = as_factor(app_holder),
         app_holder=fct_other(app_holder,keep=head(company_levels, -1),other_level = "Other")
         )%>%
  mutate(sche_type = as_factor(sch_sub_ty),
         sche_type=fct_other(sche_type,keep=c("Commercial-SAGD","Commercial-CSS"),other_level = "Other"),
         
         sche_type=fct_recode(sche_type,"SAGD"="Commercial-SAGD",
                              "CSS"="Commercial-CSS",
                              "Primary, EOR, or Other"="Other"
                              ))%>%
  mutate(label=
           ifelse(sche_name=="",paste(sche_type," (",app_holder,")",sep=""),
                  paste(sche_name," (",app_holder,")",sep="")))%>%
  I()
  


```


```{r map,fig.align='center',dpi=300,fig.width=11,fig.height=8.5}

palette<-c(colors_tableau10(),colors_tableau10_light())

ggplot(in_situ)+
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  geom_sf(aes(colour="Oil Sands Scheme Approval Boundaries",fill=app_holder),linewidth=0.5) +
  geom_label_repel(
    data=in_situ%>%filter(sche_type %in% c("CSS","SAGD")),
    aes(label = label,geometry=geometry),size=3.5, color="black",
    stat = "sf_coordinates",
    max.overlaps = Inf,
    min.segment.length = 0,
    segment.size = 0.75,         # thicker lines
    segment.colour="black",
    segment.linetype="11",
    force = 8,                     # increase repelling force between labels
  force_pull = 0.5,              # reduce pull toward anchor
  box.padding = 0.7,             # more space around labels
  )+
    #geom_sf(data=res_pts,aes(colour="Beaver Lake #131 Reserve"),fill="blue",linewidth=1.5) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(6,8,11)),size=2)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(6,  8, 11)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    #nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 2
  )+

  
  #geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
  scale_color_manual("",values=c("dodgerblue","darkgreen","black","firebrick"))+
  scale_fill_manual("Application Holder / Operator",
  values = c(
    "Harvest" = palette[1],
    "IPC" = palette[2],
    "Cenovus" = palette[3],
    "MEG" = palette[4],
    "Imperial" = palette[5],
    "Athabasca" = palette[6],
    "CNRL" = palette[7],
    "Grizzly" = palette[8],
    "Strathcona" = palette[9],
    "ConocoPhillips" = palette[10],
    "Other" = palette[11]
  )
)+
  tombe_theme()+
  guides(colour=guide_legend(nrow = 3,order = 2),fill=guide_legend(ncol = 1,order = 1))+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    legend.title = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  #facet_wrap(~year,nrow = 1)+
  labs(x="",y="",
       #title="Oil Sands Projects Within BLCN Territory",
       #subtitle="Commercial Steam-Assisted Gravity Drainage (SAGD) and Cyclic Steam Stimulation (CSS) projects are direct-labelled.\nOther projects shown are primary production or experimental regimes. Graph shows only regimes which have not expired nor been terminated or superseded.",
       #caption="Oil sands regime shapefiles via Alberta Energy Regulator ST-44, map by Andrew Leach",
       NULL)
  
ggsave("main_map.png",width=12,height = 9,dpi=300)

```


# Pipelines in BLCN Core Territory

```{r pipes,fig.align='center',fig.width=16, fig.height=8,dpi=220}

if(!file.exists("Pipeline_Inst.zip")){
  download.file("https://www.aer.ca/data/pipeline/Pipeline_Installations_SHP.zip","Pipeline_Inst.zip",mode="wb")
unzip("Pipeline_Inst.zip")
}

if(!file.exists("Pipelines_SHP.zip")){
download.file("https://www.aer.ca/data/pipeline/Pipelines_SHP.zip","Pipelines_SHP.zip",mode="wb")
unzip("Pipelines_SHP.zip")
}

#url_feat <- URLencode("https://services.arcgis.com/8df8p0NlLFEShl0r/ArcGIS/rest/services/Canada Pipelines #Correct/FeatureServer/0")


pipes2 <- st_read("Pipelines_SHP/Pipelines_GCS_NAD83.shp")%>%clean_names()%>%
  filter(substance1 %in% c("Natural Gas","Crude Oil","Hydrocarbon Diluents"))%>%
    st_transform(crs_val)

ngtl<-pipes2 %>% filter(is_neb=="Y",comp_name=="NOVA Gas Transmission Ltd.")

  
  pipes <- st_read("Pipelines_SHP/Pipelines_NAD83_10TM_AEPForest.shp")%>%clean_names()%>%
  filter(substance1 %in% c("Natural Gas","Crude Oil","Hydrocarbon Diluents"))%>%
    st_transform(crs_val)

inst_pts <- st_read("Pipeline_Installations_SHP/Pipeline_Installations_NAD83_10TM_AEPForest.shp",quiet=TRUE)%>%clean_names()%>%
  st_transform(crs_val)



inst_pts<-st_filter(inst_pts, blcn_buffer, .predicate = st_intersects)
  

#blcn_buffer<-st_buffer(blcn_terr, units::as_units(200, 'kilometer'))
oil_blcn<-st_filter(pipes2, blcn_buffer, .predicate = st_intersects)%>%
  filter(substance1=="Crude Oil")

oil_l_ath_bd<-st_filter(pipes, l_ath_bd, .predicate = st_intersects)%>%
  filter(substance1=="Crude Oil")

oil_pipes<-pipes%>%
  filter(substance1=="Crude Oil")%>%
  filter(seg_status=="Operating")


gas_pipes<-pipes%>%
  filter(substance1=="Natural Gas")#%>%
  #filter(seg_status=="Operating")


#oil_blcn<-gd_crude#%>%filter(from_fac=="Pipeline",as.numeric(out_diamet)>100)
```

```{r pipe_graph,fig.align='center',fig.width=12, fig.height=7,dpi=220}

# Create a bounding box in lon/lat (EPSG:4326)
bbox_deg <- st_as_sfc(st_bbox(c(
  xmin = -113.5,
  xmax = -108,
  ymin = 54.5,
  ymax = 56.75
), crs = 4326))

# Step 2: Transform to CRS 3347 (in metres)
bbox_m <- st_transform(bbox_deg, crs_val)

# Step 3: Extract the numeric limits in metres
bbox_coords <- st_bbox(bbox_m)


ggplot(l_ath_bd)+
  geom_sf(data=oil_pipes,aes(color="Crude Oil Pipelines"),linewidth=0.95) +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=2) +
  geom_sf(data=res_pts,aes(colour="Beaver Lake #131 Reserve"),fill="blue",linewidth=1.5) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,11,15)),size=2)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(3, 6, 7, 8, 11, 15)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 5
  )+
  scale_colour_manual("",values = c("blue","firebrick","orange"))+
  coord_sf(
    xlim = c(bbox_coords$xmin, bbox_coords$xmax),
    ylim = c(bbox_coords$ymin, bbox_coords$ymax)
  )+
  theme_minimal()+
  #scale_fill_manual("")+
 #theme_void()+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  labs(x="",y="",
      title="Crude Oil Pipeline Infrastructure Within BLCN Territory",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Alberta Energy Regulator, map by Andrew Leach")
    



```



```{r gas_pipe_graph,fig.align='center',fig.width=12, fig.height=7,dpi=220}

# Create a bounding box in lon/lat (EPSG:4326)
bbox_deg <- st_as_sfc(st_bbox(c(
  xmin = -113.5,
  xmax = -110,
  ymin = 54,
  ymax = 56.5
), crs = 4326))

# Step 2: Transform to CRS 3347 (in metres)
bbox_m <- st_transform(bbox_deg, crs_val)

# Step 3: Extract the numeric limits in metres
bbox_coords <- st_bbox(bbox_m)

point_hr <- st_as_sf(data.frame(
  lon = -112.1295892,
  lat = 55.8916186
), coords = c("lon", "lat"), crs = 4326)  # WGS84


ggplot(blcn_buffer_50)+
    geom_sf(data=gas_pipes,aes(color="Other Natural Gas Pipelines in AER Shapefile"),linewidth=0.5) +
    geom_sf(data=ngtl,aes(color="Nova Gas Transmission Line (NGTL)"),linewidth=1.15) +
	geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=2) +
  geom_sf(data=point_hr,fill=NA,size=2) +
  geom_text_repel(
    data = point_hr,
    aes(label = "House River Meter Station", geometry = geometry),
    stat = "sf_coordinates",
    nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 5
  )+
  
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake #131 Reserve"),fill="blue",linewidth=1.5) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,11,15)),size=2)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(3, 6, 7, 8, 11, 15)) %>% 
      st_transform(crs =crs_val),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 5
  )+
  scale_colour_manual("",values = c("blue","firebrick","grey10","orange"),
                      guide = guide_legend(nrow = 4))+
  coord_sf(
    xlim = c(bbox_coords$xmin, bbox_coords$xmax),
    ylim = c(bbox_coords$ymin, bbox_coords$ymax)
  )+
  theme_minimal()+
  #scale_fill_manual("")+
 #theme_void()+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  labs(x="",y="",
      title="Natural Gas Pipeline Infrastructure Within BLCN Territory",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Alberta Energy Regulator, map by Andrew Leach")
    



```






## ST-53
```{r}
#load("../st53/st53_all.rdata")

  

#exp<-in_situ %>% select(scheme_no,scheme_num,scheme_typ,approval_number,app_holder,area)
#load(file="../st53/blcn_app.rdata")
options(scipen = 999)
BLCN<-ST53_data%>%clean_names()%>%
  #for the report, data was limited to June ST-53 data
  #filter(date<=ymd("2025-05-01"))%>%
  mutate(scheme_num=str_extract(approval_number, "^\\d+"))%>%
  filter(scheme_num %in% exp$scheme_num)%>%
  mutate(project=
         if_else(is.na(scheme_name),paste(operator,recovery_method),scheme_name),
         project=gsub("Canadian Natural Resources Limited","CNRL",project),year=year(date)
         )%>%
  #scale production up to bitumen per month in m3
  mutate(monthly_prod=bitumen*days_in_month(date))%>%
  group_by(project,year)%>%
  #scale up to annual barrels
  mutate(annual_prod=sum(monthly_prod,na.rm=T))%>%
  group_by(project)%>%
  #scale up to annual barrels
  mutate(total_prod=sum(monthly_prod,na.rm=T))%>%
  ungroup()%>%
  arrange(-total_prod)%>%
  mutate(project=as_factor(project),
        )

BLCN_timing<-BLCN%>%group_by(scheme_num)%>%filter(bitumen>0)%>%
  summarize(first_date=min(date),.groups="drop")
  #save(BLCN_timing,file="ST53timing.rdata")


top_9<-levels(BLCN$project)[1:9]


BLCN<-BLCN%>%mutate(
  project=fct_other(project,keep=top_9,other_level="Other Projects")
)  %>%
  group_by(project,date,year)%>%
  #still monthly production by project (or other) in m3
  summarize(bitumen=sum(monthly_prod,na.rm = T),
            area=first(area),
            NULL)%>%
  ungroup()
```


```{r st53_graph,fig.align='center',fig.width=12, fig.height=7,dpi=220}
BLCN%>%
  group_by(project,year)%>%
  #still monthly production by project (or other) in m3
  summarize(bitumen=sum(bitumen,na.rm = T)/10^6*6.2898)%>%
  ggplot(aes(year,bitumen))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  scale_x_continuous(expand=c(0,0),breaks=pretty_breaks(6))+
  scale_y_continuous(expand=c(0,0),breaks=pretty_breaks(5))+
  #coord_flip()+
  #scale_fill_brewer()+
  #guides(fill="none",color="none")+
  facet_wrap(~project,nrow=2)+
  #scale_x_reverse()+
  theme_ps_grid()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Annual Bitumen Production (million barrels)",
       title="Annual Bitumen Production by Project, Beaver Lake Cree Nation Core Territory",
       caption="Data via Alberta Energy Regulator ST-53, graph by Andrew Leach")  
  
```

```{r st53_graph_cumul,fig.align='center',fig.width=12, fig.height=7,dpi=220}
#test<-
  BLCN%>%
  group_by(project)%>%
  #still monthly production by project (or other) in m3
  mutate(bitumen_cumul=cumsum(bitumen)/10^9*6.2898)%>% ungroup()%>%
  ggplot(aes(date,bitumen_cumul))+
  geom_area(aes(fill=project),size=0.25,position = "stack",color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  scale_fill_manual("",values = rev(colors_colorblind_10()))+
  scale_x_date(expand=c(0,0),breaks=pretty_breaks(6))+
  scale_y_continuous(expand=c(0,0),breaks=pretty_breaks(5))+
  #scale_x_reverse()+
  theme_ps_grid()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5),
        legend.position = "right"
        )+
  labs(x=NULL,y="Cumulative Bitumen Production (billion barrels)",
       title="Cumulative Bitumen Production by Project, Beaver Lake Cree Nation Core Territory",
       caption="Data via Alberta Energy Regulator ST-53, calculations and graph by Andrew Leach")  
  
```


```{r st53_graph_2,fig.align='center',fig.width=10, fig.height=5,dpi=500}
# Calculate cumulative and last value per project
BLCN_labels <- 
  BLCN %>%
  group_by(project) %>%
  mutate(bitumen_cumul = cumsum(bitumen) / 10^9 * 6.2898) %>%
  filter(date == max(date)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(bitumen_cumul=cumsum(bitumen_cumul))%>%arrange(n)%>%
  mutate(lead=lead(bitumen_cumul),lead = replace_na(lead, 0))%>%
  mutate(bitumen_cumul=bitumen_cumul-(bitumen_cumul-lead)/2)
  

# Main plot
#cumul_prod_project<-
BLCN %>%
  group_by(project) %>%
  mutate(bitumen_cumul = cumsum(bitumen) / 10^9 * 6.2898) %>%
  ungroup() %>%
  ggplot(aes(date, bitumen_cumul)) +
  geom_area(aes(fill = project), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = date+months(1), y = bitumen_cumul, label = project),
    hjust = 0, size = 3, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_x_date(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_ps_grid() +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 100, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Cumulative Bitumen Production (billion barrels)",
    #title = "Cumulative Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    caption = "Data via Alberta Energy Regulator ST-53, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_prod_project,file="cumulative_prod_project.png",dpi=500,bg="white",width=10,height=5)

```

```{r st53_graph_2,fig.align='center',fig.width=8, fig.height=5,dpi=300}
# Calculate cumulative and last value per project
BLCN_labels <- 
  BLCN %>%
  filter(date == max(date)) %>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(bitumen_cumul=cumsum(bitumen* 6.2898/days_in_month(date)/10^3))%>%
  mutate(lead=lead(bitumen_cumul),lead = replace_na(lead, 0))%>%
  mutate(lag=lag(bitumen_cumul),lag = replace_na(lag, 0))%>%
  mutate(label=lag+(bitumen_cumul-lag)/2)%>%
  I()

# Main plot
monthly_prod_project<-
BLCN %>%
  ggplot(aes(date, bitumen* 6.2898/days_in_month(date)/10^3)) +
  geom_area(aes(fill = project), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = date+months(1), y = label, label = project),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_x_date(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
  expand_limits(x=ymd("2000-01-01"))+
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Monthly Bitumen Production (thousand barrels per day)",
    #title = "Cumulative Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Alberta Energy Regulator ST-53, calculations and graph by Andrew Leach"
  )

ggsave(monthly_prod_project,file="monthly_prod_project.png",dpi=300,bg="white",width=12,height=5.5)

monthly_prod_project
```


### Annual Production (mmbbl/d)
```{r, echo = T,eval=T}
  BLCN%>% group_by(date,year)%>%
  #still monthly production by project (or other) in m3
  summarize(bitumen=sum(bitumen,na.rm = T)/10^6*6.2898,.groups="drop")%>%
  mutate(days=days_in_month(date))%>%
  group_by(year)%>%
  summarize(bitumen=sum(bitumen),days=sum(days),bitumen=bitumen/days)%>%
  select(Year=year,"Production (mmbbl/d)"=bitumen)%>%
  mutate(Year=as.character(Year),
         #Year=ifelse(Year=="2025","2025 (to date)",Year)
         )%>%
   kbl(escape = FALSE,table.attr = "style='width:60%;'") %>%
    kable_styling(fixed_thead = T,bootstrap_options = c("hover", "condensed","responsive"),full_width = T)%>%
  #scroll_box(width = "1000px", height = "600px")%>%
  column_spec(1, extra_css = "white-space: nowrap;") %>%
      I() 
```




```{r,eval=F}

#Sproule Valuation
# Load the Sproule price forecast page
url <- "https://sproule.com/price-forecast/"
page <- read_html(url)

# Extract all links
all_links <- page %>%
  html_elements("a") %>%
  html_attr("href") %>%
  na.omit()

# Grab Forecast Excel Files
forecast_links <- all_links %>%
  str_subset("(?i)escalated.*\\.xlsx$")  #  .xlsx files containing "escalated"

download.file(first(forecast_links),destfile = "sproule_current.xlsx",mode="wb")


```


```{r echo=F,message=F,warning=F,include=F}

download.file("https://sproule-erce.com/wp-content/uploads/2026/03/2026-02-Escalated.xlsx",destfile = "sproule_feb.xlsx",mode="wb")

#sproule_hist<-read_excel("sproule_current.xlsx",sheet=" History",skip=9)%>%
  sproule_hist<-read_excel("sproule_feb.xlsx",sheet=" History",skip=9)%>%
  clean_names()%>%
  mutate(month=excel_numeric_to_date(as.numeric(month)))%>%
  filter(!is.na(month))%>%
  rename(date=month)%>%
  mutate(across(-date, ~ as.numeric(.)))%>%
  filter(!if_all(-date, is.na))

header_rows <- read_excel("sproule_current.xlsx",sheet=" History",skip=6,col_names = FALSE)%>%
  slice(1:3)


header_rows <- read_excel("sproule_current.xlsx",
                          sheet = " History",
                          skip = 6,
                          n_max = 3,      # Only get 3 rows
                          col_names = FALSE)




# Check it's actually a data frame
if (!is.data.frame(header_rows)) stop("Header rows not read correctly.")
#create names

collapsed_names <- map_chr(seq_len(ncol(header_rows)), function(i) {
  parts <- as.character(header_rows[[i]])
  parts <- parts[!is.na(parts)]
  if (length(parts) == 0) {
    return(paste0("V", i))  # fallback name if all NA
  } else {
    return(paste(parts, collapse = "_"))
  }
}) %>%
  str_replace_all("_+", "_") %>%
  str_remove("^_") %>%
  str_remove("\\*+") %>%
  str_remove("_$")%>%
  str_replace_all("_"," ")
  


collapsed_names<-c("date",collapsed_names)
# Apply the new names
names(sproule_hist) <- make.unique(collapsed_names)


unit_row <- read_excel("sproule_current.xlsx",
                          sheet = " History",
                          skip = 9,
                          n_max = 1,      # Only get 1 rows
                          col_names = FALSE)%>%
  as.character()

# Create the 2-column tibble
name_unit_df <- tibble(
  commodity = collapsed_names,
  unit = unit_row
)%>%mutate(commodity=make.unique(commodity))



sproule_hist<-sproule_hist %>% pivot_longer(-date,names_to = "commodity",values_to = "price")%>%left_join(name_unit_df,relationship = "many-to-many")%>%
  filter(!is.na(price))

```



```{r}
#get cpi data to adjust to real
#cpi<-cansim::get_cansim("18-10-0004-01")
base_year<-2025
cpi<-cansim::get_cansim_vector('v41692327') %>% clean_names()%>%
  mutate(date=ymd(ref_date))%>%
  mutate(index_val=value/sum(value*(year(date)==2025)/12)*100)%>%
  ungroup()%>%
  select(date,geo,cpi_2025=index_val,cpi_2002=value)
```


```{r}
diffs<- sproule_hist%>%filter(commodity%in%c("WTI Cushing Oklahoma","Edmonton Par Price 40 API","Brent UK 38 API",             "Hardisty Heavy 12 API","Exchange Rate","SCO at Edmonton 32 API","Cromer LSB 35 API","Canadian Light Sweet","Louisiana Light Sweet",                                     "Edmonton Condensate","WCS 20.5 API","Cold Lake Blend 22.6 API" 
                                           ),
                                  unit %in%c("$/bbl","$US/bbl","CAD/USD"))%>%
                                  select(-unit)%>%
                                  pivot_wider(names_from = commodity,values_from = price)%>%
                                  mutate("Edmonton Light Sweet"=ifelse(is.na(`Edmonton Par Price 40 API`),`Canadian Light Sweet`,`Edmonton Par Price 40 API`))%>%
                                  rename("cad_usd"="Exchange Rate")%>%
                                  rename("Brent (UK)"="Brent UK 38 API")%>%
                                  rename("Hardisty Heavy Sour"="Hardisty Heavy 12 API")%>%
                                  mutate(`Edmonton Light Sweet ($US)`=`Edmonton Light Sweet`*cad_usd)%>%
                                  mutate(`Hardisty Heavy Sour ($US)`=`Hardisty Heavy Sour`*cad_usd)%>%
                                  select(-c(`Edmonton Par Price 40 API`,`Canadian Light Sweet`))%>%
  pivot_longer(-c(date,cad_usd),names_to = "commodity",values_to = "price")


diffs_plot<-diffs %>% filter(commodity %in% c("Brent (UK)","WTI Cushing Oklahoma","Edmonton Light Sweet ($US)","Hardisty Heavy Sour ($US)"))%>%
  mutate(commodity=as_factor(commodity),
         commodity=fct_relevel(commodity,"Edmonton Light Sweet ($US)"))%>%
  pivot_wider(values_from = price,names_from = commodity)%>%
  mutate(`Edmonton Light Sweet vs. WTI`=`Edmonton Light Sweet ($US)`-`WTI Cushing Oklahoma`)%>%
  mutate(`Brent (UK) vs. WTI`=`Brent (UK)`-`WTI Cushing Oklahoma`)%>%
  mutate(`Edmonton Light Sweet vs. Hardisty Heavy Sour`=`Edmonton Light Sweet ($US)`-`Hardisty Heavy Sour ($US)`)%>%
  select(1,7,8,9)%>%
  pivot_longer(-date,values_to = "diff",names_to="commodity")%>%
  mutate(commodity=as_factor(commodity))%>%
  ggplot() +
  geom_line(aes(date,diff,group = commodity,colour=commodity,lty=commodity),linewidth=0.85) +
  scale_colour_manual("",values = c("black","grey60","black"))+  # or "Signac" "Paquin"
  scale_linetype_manual("",values = c("solid","11","21"))+  # or "Signac" "Paquin"
  scale_x_date(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
  expand_limits(y=0)+
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_ps_grid() +
  labs(
    x = NULL,
    y = "Difference between Monthly Price ($US/barrel)",
    #title = "Cumulative Estimated Value of Bitumen Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Alberta Energy Regulator ST-53 and Sproule, calculations and graph by Andrew Leach"
  )

diffs_plot<-diffs_plot +
  theme_irpp() +
  guides(colour=guide_legend(nrow = 3))+
  geom_hline(yintercept = 0, color = "black", linewidth = 0.4) +
  theme(
    #axis.line = element_line(color = "black", linewidth = 0.4),
    #axis.ticks = element_line(color = "black", linewidth = 0.4),
    
    #plot.title = element_text(size = 10, color = "black"),
    #axis.title  = element_text(size = 9, color = "black"),
    
    # x-axis labels shifted slightly down from ticks
    #text = element_text(size = 12, color = "black"),
    axis.text.x = element_text(angle=90),
    
    # y-axis labels unchanged
    #axis.text.y = element_text(size = 12, color = "black"),
    
    #panel.grid = element_blank(),
    #legend.position = "bottom",
    #plot.margin = margin(t = 5, r = 30, b = 5, l = 5)
  )
ggsave(plot = diffs_plot,"diffs.png",bg="white",width = 5,height=5,dpi=300)
ggsave(plot = diffs_plot,"diffs.svg",bg="white",width = 5,height=5,dpi=300)



```



```{r}
BLCN_val<-BLCN %>%
  #filter(year>=2008)%>% #for 2008-2025 calcs
  mutate(date=floor_date(date,"month"))%>%
  left_join(
    #test<-
      sproule_hist%>%filter(commodity%in%c("Hardisty Heavy 12 API","WTI Cushing Oklahoma","Exchange Rate",
                                       "Edmonton Condensate","WCS 20.5 API","Cold Lake Blend 22.6 API" 
                                           ),
                                  unit %in%c("$/bbl","$US/bbl","CAD/USD"))%>%
                                  select(-unit)%>%
                                  pivot_wider(names_from = commodity,values_from = price)%>%
                                  clean_names()%>%rename("cad_usd"=exchange_rate)
  )%>%
  left_join(cpi)%>%
  
  mutate(wti_cad=wti_cushing_oklahoma/cad_usd,
        month=month(date),
         diluent_share = case_when(
  month == 1  ~ 0.359504931,
  month == 2  ~ 0.360247247,
  month == 3  ~ 0.35689778,
  month == 4  ~ 0.356666451,
  month == 5  ~ 0.337556682,
  month == 6  ~ 0.32928475,
  month == 7  ~ 0.314181266,
  month == 8  ~ 0.30623779,
  month == 9  ~ 0.315789474,
  month == 10 ~ 0.333266278,
  month == 11 ~ 0.338327341,
  month == 12 ~ 0.357055056 # based on BVM shares for 2024
))%>%
  mutate(
  gross_rev_royalty = case_when(
  wti_cad<55  ~ 0.01,
  wti_cad>120 ~ 0.09,
  TRUE ~ 0.01+.08/65*(wti_cad-55)
),
  gross_rev_royalty = case_when(
  project=="Cold Lake"  ~ 0.05, #cold lake framework
  
  TRUE ~ gross_rev_royalty
),



# based on BVM shares for 2024
         )%>%
  mutate(bvm_val=(wcs_20_5_api-diluent_share*edmonton_condensate)/(1-diluent_share),
         bit_val=ifelse(is.na(bvm_val),hardisty_heavy_12_api,bvm_val),
         #bit_val=hardisty_heavy_12_api,
         transp_cost=.15*(area!="Athabasca")+.92*(area=="Athabasca"),
         transp_cost=transp_cost/100*cpi_2025,#convert to real
         bit_val=bit_val-transp_cost-6.32/100*cpi_2025,#real dollar discount for transport and quality. Choose to align w royalty values
         bit_val=pmax(bit_val,0)
         )%>%
  
  #convert monthly production to barrels and multiply by Hardisty Heavy Price
    #adjusting for $45/m3 transportation and quality diff to Hardisty Heavy
  group_by(project)%>%
  mutate(bitumen_bbl=bitumen*6.2898,
         value=bitumen_bbl*bit_val,
         roy_val=value*gross_rev_royalty*(year(date)>2008)+value*0.01*(year(date)==2008), #calculate royalty from 2008
         value_cumul=cumsum(value),
         royalty_cumul=cumsum(roy_val),
         cumul_bitumen=cumsum(bitumen_bbl),
         value_bbl_avg=value_cumul/cumul_bitumen,
         roy_value_bbl_avg=royalty_cumul/cumul_bitumen,
         value_2025=value/(cpi_2025/100),
         real_cumul=cumsum(value_2025),
         roy_val_2025=roy_val/(cpi_2025/100),
         real_roy_cumul=cumsum(roy_val_2025)
                            )%>%ungroup()


```

```{r}
real<-sproule_hist %>% left_join(cpi)%>%
  mutate(real=price/(cpi_2025/100))

real %>% filter(commodity=="WTI Cushing Oklahoma")%>%
  ggplot()+geom_line(aes(date,price,colour="Nominal"))+
  geom_line(aes(date,real,colour="Real"))

```


```{r st53val_graph,fig.align='center',fig.width=12, fig.height=7,dpi=220}

value_labels <- BLCN_val %>%
  filter(date == max(date)) %>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(value_cumul=cumsum(value_cumul))%>%
  select(n,date,project,value_cumul)%>%
  arrange(n)%>%
  mutate(lead=lead(value_cumul),lead = replace_na(lead, 0))%>%
  mutate(label_val=value_cumul-(value_cumul-lead)/2,
         label_val=label_val/10^9)

cumul_val_project<-
  
BLCN_val %>%
  ggplot(aes(date, value_cumul/10^9)) +
  geom_area(aes(fill = project), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = value_labels,
    aes(x = date+months(1), y = label_val, label = project),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_x_date(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  expand_limits(y=300)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Estimated Value of Bitumen Production (billion dollars)",
    #title = "Cumulative Estimated Value of Bitumen Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Alberta Energy Regulator ST-53 and Sproule, calculations and graph by Andrew Leach"
  )

ggsave(cumul_val_project,file="cumulative_prod_value.png",dpi=300,bg="white",width=10,height=5)
cumul_val_project


```


```{r}
#basic calcs
BLCN_val %>%
  filter(year<2016)%>%
  filter(year>=2008)%>%
  #filter(year<=2008)%>%
  summarize(prod=sum(bitumen_bbl)/10^9,
            royalty=sum(roy_val)/10^9,
            roy_per_bbl=royalty/prod,
            royalty_real=sum(roy_val_2025)/10^9,
            roy_real_per_bbl=royalty_real/prod,
            value=sum(value)/10^9,
            real_value=sum(value_2025)/10^9,
            )

royalties_old<-
  BLCN_val %>%
  filter(year(date)<2016)%>%
  filter(year(date)>=2008)%>%
  group_by(year)%>%
  summarize(prod=sum(bitumen_bbl)/10^9,
            royalty=sum(roy_val)/10^9,
            roy_per_bbl=royalty/prod,
            royalty_real=sum(roy_val_2025)/10^9,
            roy_real_per_bbl=royalty_real/prod,
            value=sum(value)/10^9,
            real_value=sum(value_2025)/10^9,
            )

```



```{r st53val_graph_real,fig.align='center',fig.width=12, fig.height=7,dpi=220}

value_labels <- BLCN_val %>%
  filter(date == max(date)) %>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(value_cumul=cumsum(real_cumul))%>%
  select(n,date,project,value_cumul)%>%
  arrange(n)%>%
  mutate(lead=lead(value_cumul),lead = replace_na(lead, 0))%>%
  mutate(label_val=value_cumul-(value_cumul-lead)/2,
         label_val=label_val/10^9)

real_val_project<-
  
BLCN_val %>%
  ggplot(aes(date, real_cumul/10^9)) +
  geom_area(aes(fill = project), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = value_labels,
    aes(x = date+months(1), y = label_val, label = project),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_x_date(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  expand_limits(y=300)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Estimated Value of Bitumen Production (billion real ($2025) dollars)",
    #title = "Cumulative Estimated Value of Bitumen Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Alberta Energy Regulator ST-53 and Sproule, calculations and graph by Andrew Leach"
  )
real_val_project
ggsave(real_val_project,file="cumulative_prod_value.png",dpi=300,bg="white",width=10,height=5)



```


```{r st53_roy_val_graph,fig.align='center',fig.width=12, fig.height=7,dpi=220}

value_labels <- BLCN_val %>%
  filter(year(date)<2016)%>%
  filter(year(date)>=2008)%>%
  filter(date == max(date)) %>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(royalty_cumul=cumsum(royalty_cumul))%>%
  select(n,date,project,royalty_cumul)%>%
  arrange(n)%>%
  mutate(lead=lead(royalty_cumul),lead = replace_na(lead, 0))%>%
  mutate(label_val=royalty_cumul-(royalty_cumul-lead)/2,
         label_val=label_val/10^9)

#cumul_val_project<-
  
BLCN_val %>%
  filter(year(date)<2016)%>%
  filter(year(date)>=2008)%>%
  #mutate(year=year(date))%>%
  #group_by(year,project)%>%
  ggplot(aes(date, royalty_cumul/10^9)) +
  geom_area(aes(fill = project), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = value_labels,
    aes(x = date+months(1), y = label_val, label = project),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_x_date(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  expand_limits(y=5)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Estimated Minimum Cumulative Value of Bitumen Royalties (billion dollars)",
    #title = "Cumulative Estimated Value of Bitumen Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Alberta Energy Regulator ST-53 and Sproule, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_val_project,file="cumulative_prod_value.png",dpi=300,bg="white",width=10,height=5)



```

```{r}


BLCN_val %>%
  filter(year(date)<2017)%>%
  filter(year(date)>=2008)%>%
  mutate(year=year(date))%>%
  group_by(year,project)%>%
  summarize(roy_val=sum(roy_val))%>%
  ggplot(aes(year, roy_val/10^9)) +
  geom_col(aes(fill = project), size = 0.25, position = "stack", color = "black") +
  #geom_text(
  #  data = value_labels,
  #  aes(x = date+months(1), y = label_val, label = project),
  #  hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  #) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_x_continuous(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  #expand_limits(y=5)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "right", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Estimated Minimum Cumulative Value of Bitumen Royalties (billion dollars)",
    #title = "Cumulative Estimated Value of Bitumen Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Alberta Energy Regulator ST-53 and Sproule, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_val_project,file="cumulative_prod_value.png",dpi=300,bg="white",width=10,height=5)



```

```{r st53_roy_val_graph_real,fig.align='center',fig.width=12, fig.height=7,dpi=220}

value_labels <- BLCN_val %>%
  filter(year(date)<2016)%>%
  filter(year(date)>=2008)%>%
  filter(date == max(date)) %>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(royalty_cumul=cumsum(real_roy_cumul))%>%
  select(n,date,project,royalty_cumul)%>%
  arrange(n)%>%
  mutate(lead=lead(royalty_cumul),lead = replace_na(lead, 0))%>%
  mutate(label_val=royalty_cumul-(royalty_cumul-lead)/2,
         label_val=label_val/10^9)

#cumul_val_project<-
  
BLCN_val %>%
  filter(year(date)<2016)%>%
  filter(year(date)>=2008)%>%
  #mutate(year=year(date))%>%
  #group_by(year,project)%>%
  ggplot(aes(date, real_roy_cumul/10^9)) +
  geom_area(aes(fill = project), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = value_labels,
    aes(x = date+months(1), y = label_val, label = project),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_x_date(
  expand = c(0, 0),
  breaks = pretty_breaks(12),
  labels = date_format("%Y")
) +
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  expand_limits(y=5)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Estimated Minimum Cumulative Value of Bitumen Royalties (billion 2025 real dollars)",
    #title = "Cumulative Estimated Value of Bitumen Production by Project (2000-2025), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Alberta Energy Regulator ST-53 and Sproule, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_val_project,file="cumulative_prod_value.png",dpi=300,bg="white",width=10,height=5)



```


## Royalty Data

All data on royalties for 2016-2024 are posted by the Government of Alberta [here](https://open.alberta.ca/opendata/alberta-oil-sands-royalty-data1). I have compiled each of the annual files into a single spreadsheet for use for my report. The code below processes the royalty data for use in my expert report for *Beaver Lake Cree Nation et al v. His Majesty the King in Right of the Province of Alberta, et al.*, Court File No.: 0803 06718, Alberta Court of King's Bench. 

<!-- ## Download and Combine the Annual Data for 2016-2022 -->

```{r load data,cache=T,include=F}
#load plant data
make_os_data<-function(){
os_data_2024 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2024 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2023 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2023 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2022 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2022 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
os_data_2021 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2021 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2020 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2020 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2019 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2019 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  os_data_2018 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2018 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2017 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2017 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
os_data_2016 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2016 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

#stack all the years together

os_data<-bind_rows(os_data_2024,os_data_2023,os_data_2022,os_data_2021,os_data_2020,os_data_2019,os_data_2018,os_data_2017,os_data_2016)

#clean the names

os_data<-os_data %>% clean_names()%>%
  mutate(op_profit=gross_revenue-operating_costs-other_costs+other_net_proceeds)

# data are reported for an individual project twice if it passes payout in a given year
# combine the data and include two things: a payout year indicator and a net and gross rev royalty number along with a total royalty paid
os_data <- os_data %>% group_by(project,reporting_year,project_name,operator_name)%>%
  summarize(
    project_revenue=sum(project_revenue),
    gross_revenue=sum(gross_revenue),
    op_profit=sum(op_profit),
    cleaned_crude_bitumen_at_rcp_barrels=sum(cleaned_crude_bitumen_at_rcp_barrels),
    operating_costs=sum(operating_costs),
    capital_costs=sum(capital_costs),
    return_allowance=sum(return_allowance),
    other_costs=sum(other_costs),
    other_net_proceeds=sum(other_net_proceeds),
    net_revenue=sum(net_revenue),
    royalty_type=last(royalty_type),
    royalty_type_start=first(royalty_type),
    payout_start=first(payout_status),
    payout_status=last(payout_status),
    net_rev_royalty=sum(royalty_payable*(royalty_type=="NET")),
    gross_rev_royalty=sum(royalty_payable*(royalty_type=="GROSS")),
    royalty_payable=sum(royalty_payable),
    royalty_rate=last(royalty_rate_percent),
    first_royalty_rate=first(royalty_rate_percent),
    royalty_avg=royalty_payable/gross_revenue,
    payout_year=(n()>1),
    unrecovered_balance_net_loss_at_eop=last(unrecovered_balance_net_loss_at_eop))%>%
    mutate(
      last_prod=last(cleaned_crude_bitumen_at_rcp_barrels)/365,
      label=paste(project_name,"\n(",formatC(round(last_prod,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""))%>%
  ungroup()%>%
  group_by(project)%>%
  mutate(project_payout=(last(payout_status)=="POST"))

#fix facility names
os_data <- os_data %>% mutate(
  royalty_bbl=royalty_payable/cleaned_crude_bitumen_at_rcp_barrels,
  op_costs_bbl=operating_costs/cleaned_crude_bitumen_at_rcp_barrels,
  gross_revenue_bbl=gross_revenue/cleaned_crude_bitumen_at_rcp_barrels,
  cap_costs_bbl=capital_costs/cleaned_crude_bitumen_at_rcp_barrels,
  project_name=gsub(" Project","",project_name),
  project_name=gsub("Christina Lake Regional","Christina Lake (MEG)",project_name),
  project_name=gsub("Christina Lake Thermal","Christina Lake (CVE)",project_name),
  project_name=gsub("MacKay River Commercial","PetroChina",project_name),
  project_name=gsub("MacKay River","MacKay River (Suncor)",project_name),
  project_name=gsub("PetroChina","MacKay River (PetroChina)",project_name),
  project_name=gsub("Harvest BlackGold","BlackGold",project_name),
  
  project_name=gsub(" Thermal","",project_name),
  project_name=gsub(" Mine","",project_name),
  project_name=gsub(" Oil Sands","",project_name),
  project_name=gsub(" EOR","",project_name),
  project_name=gsub(" Commercial","",project_name),
  project_name=gsub(" SAGD","",project_name),
  project_name=gsub(" Demonstration","",project_name),
  project_name=gsub(" In-Situ","",project_name),
  project_name=as.factor(project_name),
  op_profit_net=op_profit-royalty_payable,
  op_profit_bbl=op_profit/cleaned_crude_bitumen_at_rcp_barrels,
  op_profit_net_bbl=op_profit_net/cleaned_crude_bitumen_at_rcp_barrels
  )

}

os_data<-make_os_data()

os_data<-os_data %>% mutate(blcn=(project %in% os_project_map$ProjectID),
                            )%>%
  group_by(project_name)%>%
  mutate(total_bbls=cumsum(cleaned_crude_bitumen_at_rcp_barrels),
         min_bbls=min(cleaned_crude_bitumen_at_rcp_barrels)/365,
         max_bbls=max(cleaned_crude_bitumen_at_rcp_barrels)/365)%>%
  ungroup()

blcn_data<-
      os_data %>% filter(blcn)%>%
        arrange(-total_bbls)%>%
         mutate(project_name=as.character(project_name),
                project_name=as_factor(project_name)
                ) %>% 
  #join annual average cpi to adjust to 2025 dollars
  left_join(cpi %>% mutate(reporting_year=year(date))%>%
                                  group_by(reporting_year)%>%summarize(cpi_2025=mean(cpi_2025)))

  
#make spreadsheet file for report writing
write.csv(file = "projects.csv",blcn_data%>%select(project,project_name,operator_name)%>%
            group_by(project)%>%
            summarize(project_name=last(str_to_title(project_name)),operator_name=last(str_to_title(operator_name))))


#save the data file
save(blcn_data,file = "os_data_blcn.rdata")

#summarize the total for these projects
total<-blcn_data %>% 
         summarize(royalties=sum(royalty_payable,na.rm=T)/10^9,
                   royalties_real=sum(royalty_payable/(cpi_2025/100),na.rm=T)/10^9,
                   rev=sum(gross_revenue,na.rm=T)/10^9,
                   rev_real=sum(gross_revenue/(cpi_2025/100),na.rm=T)/10^9,
                   prod=sum(cleaned_crude_bitumen_at_rcp_barrels,na.rm=T)/10^9,
                   rev_per_bbl=rev/prod,
                   rev_real_per_bbl=rev_real/prod
                   
                   )

royalties_new<-
  blcn_data %>%
  rename(year=reporting_year)%>%
  group_by(year)%>%
     summarize(royalty=sum(royalty_payable,na.rm=T)/10^9,
                   royalty_real=sum(royalty_payable/(cpi_2025/100),na.rm=T)/10^9,
                   value=sum(gross_revenue,na.rm=T)/10^9,
                   real_value=sum(gross_revenue/(cpi_2025/100),na.rm=T)/10^9,
                   prod=sum(cleaned_crude_bitumen_at_rcp_barrels,na.rm=T)/10^9,
                   roy_per_bbl=royalty/prod,
                   roy_real_per_bbl=royalty_real/prod)


#royalties<-
  royalties_old %>% bind_rows(royalties_new)%>%
  mutate(disc=2025-year)%>%
  summarize(prod=sum(prod),
            nominal=sum(royalty)+5.2,
            real=sum(royalty_real)+5.2,
            npv_r5=sum(royalty_real/(1.05^disc))+5.2,
            npv_r7=sum(royalty_real/(1.07^disc))+5.2,
            npv_r10=sum(royalty_real/(1.1^disc))+5.2,
            npv_r15=sum(royalty_real/(1.15^disc))+5.2
            )

```



```{r}
#combine projects keeping only the largest 9 and capturing other projects in one category
project_levels <- c(Jackfish = "Jackfish",Jackfish = "Jackfish 2",Jackfish = "Jackfish 3",Kirby = "Kirby North",Kirby = "Kirby South")


blcn_data<-
  #test<-
  blcn_data%>%
        mutate(
          project_name=fct_recode(project_name, !!!project_levels),
          project_name=fct_other(project_name,keep=c(head(levels(blcn_data$project_name),5),"Jackfish","Kirby"),other_level = "Other Projects"))%>%
        group_by(project_name,reporting_year)%>%
        summarize(prod=sum(cleaned_crude_bitumen_at_rcp_barrels,na.rm = T),
                  royalties=sum(royalty_payable,na.rm = T),
                  revenues=sum(gross_revenue,na.rm=T),
                  rev_real=sum(gross_revenue/cpi_2025*100,na.rm=T),
                  royalties_real=sum(royalty_payable/cpi_2025*100,na.rm=T)
                    )%>%
  ungroup()

```

```{r prod_by_project,fig.align='center',fig.width=10, fig.height=5,dpi=500}

# Calculate cumulative and last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(bitumen_cumul = cumsum(prod)/10^9,
         year=as.numeric(reporting_year)) %>%
  filter(year == max(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(bitumen_cumul=cumsum(bitumen_cumul))%>%arrange(n)%>%
  mutate(lead=lead(bitumen_cumul),lead = replace_na(lead, 0))%>%
  mutate(bitumen_cumul=bitumen_cumul-(bitumen_cumul-lead)/2)
  

# Main plot
#cumul_prod_project<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(bitumen_cumul = cumsum(prod) / 10^9,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, bitumen_cumul)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = bitumen_cumul, label = project_name),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Cumulative Bitumen Production (billion barrels)",
    #title = "Cumulative Production by Royalty Project (2016-2023), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_prod_project,file="cumulative_prod_project.png",dpi=500,bg="white",width=10,height=5)

```





```{r annual_prod_by_project,fig.align='center',fig.width=10, fig.height=5,dpi=500}

# Calculate cumulative and last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(prod=prod/10^6,
         year=as.numeric(reporting_year)) %>%
  filter(year == max(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(prod=cumsum(prod))%>%arrange(n)%>%
  mutate(lead=lead(prod),lead = replace_na(lead, 0))%>%
  mutate(prod=prod-(prod-lead)/2)
  

# Main plot
#cumul_prod_project<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(prod=(prod) / 10^6,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, prod)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = prod, label = project_name),
    hjust = 0, size = 3, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_ps_grid() +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 100, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Annual Bitumen Production (million barrels)",
    #title = "Annual Production by Royalty Project (2016-2023), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_prod_project,file="cumulative_prod_project.png",dpi=500,bg="white",width=10,height=5)

```




```{r royalty_by_project,fig.align='center',fig.width=10, fig.height=6,dpi=500}

# Calculate cumulative and last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(royalty_cumul = cumsum(royalties)/10^9,
         year=as.numeric(reporting_year)) %>%
  filter(year == max(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(royalty_cumul=cumsum(royalty_cumul))%>%arrange(n)%>%
  mutate(lead=lead(royalty_cumul),lead = replace_na(lead, 0))%>%
  mutate(royalty_cumul=royalty_cumul-(royalty_cumul-lead)/2)
  

# Main plot
#cumul_prod_project<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(royalty_cumul = cumsum(royalties) / 10^9,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, royalty_cumul)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = royalty_cumul, label = project_name),
    hjust = 0, size = 3, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_ps_grid() +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 100, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Cumulative Royalty Payments (billion dollars)",
    title = "Cumulative Royalty Payments by Royalty Project (2016-2024), Beaver Lake Cree Nation Core Territory",
    caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_prod_project,file="cumulative_prod_project.png",dpi=500,bg="white",width=10,height=5)

```

```{r real_royalty_by_project,fig.align='center',fig.width=10, fig.height=6,dpi=500}

# Calculate cumulative and last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(royalty_cumul = cumsum(royalties_real)/10^9,
         year=as.numeric(reporting_year)) %>%
  filter(year == max(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(royalty_cumul=cumsum(royalty_cumul))%>%arrange(n)%>%
  mutate(lead=lead(royalty_cumul),lead = replace_na(lead, 0))%>%
  mutate(royalty_cumul=royalty_cumul-(royalty_cumul-lead)/2)
  

# Main plot
#cumul_prod_project<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(royalty_cumul = cumsum(royalties_real) / 10^9,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, royalty_cumul)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = royalty_cumul, label = project_name),
    hjust = 0, size = 3, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_ps_grid() +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 100, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Cumulative Royalty Payments (billion 2025 real dollars)",
    title = "Cumulative Royalty Payments by Royalty Project (2016-2024), Beaver Lake Cree Nation Core Territory",
    caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_prod_project,file="cumulative_prod_project.png",dpi=500,bg="white",width=10,height=5)

```


```{r annual_royalty_by_project,fig.align='center',fig.width=8, fig.height=4,dpi=500}

# Calculate cumulative and last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(royalties=(royalties)/10^9,
         year=as.numeric(reporting_year)) %>%
  filter(year == last(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(royalties=cumsum(royalties))%>%arrange(n)%>%
  mutate(lead=lead(royalties),lead = replace_na(lead, 0))%>%
  mutate(royalties=royalties-(royalties-lead)/2)
  

# Main plot
#royalties_project<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(royalties =(royalties) / 10^9,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, royalties)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = royalties, label = project_name),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(8)) +
  expand_limits(y=8)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Annual Royalty Payments (billion dollars)",
    #title = "Annual Royalty Payments by Royalty Project (2016-2023), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(royalties_project,file="royalties_project.png",dpi=500,bg="white",width=10,height=5)

```

```{r cumul_rev,fig.align='center',fig.width=10, fig.height=6,dpi=500}

# Calculate cumulative and last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(rev_cumul = cumsum(revenues)/10^9,
         year=as.numeric(reporting_year)) %>%
  filter(year == max(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(rev_cumul=cumsum(rev_cumul))%>%arrange(n)%>%
  mutate(lead=lead(rev_cumul),lead = replace_na(lead, 0))%>%
  mutate(rev_cumul=rev_cumul-(rev_cumul-lead)/2)
  

#sum(blcn_data$revenues)/10^9
# Main plot
#cumul_rev_project<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(rev_cumul = cumsum(revenues) / 10^9,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, rev_cumul)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = rev_cumul, label = project_name),
    hjust = 0, size = 3, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(5)) +
  theme_ps_grid() +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 100, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Cumulative Revenue (billion dollars)",
    title = "Cumulative Project Revenue by Royalty Project (2016-2024), Beaver Lake Cree Nation Core Territory",
    caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(cumul_prod_project,file="cumulative_prod_project.png",dpi=500,bg="white",width=10,height=5)

```



```{r annual_rev_by_project,fig.align='center',fig.width=12, fig.height=5.5,dpi=500}

# Calculate last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(revenues=(revenues)/10^9,
         year=as.numeric(reporting_year)) %>%
  filter(year == max(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(revenues=cumsum(revenues))%>%arrange(n)%>%
  mutate(lead=lead(revenues),lead = replace_na(lead, 0))%>%
  mutate(revenues=revenues-(revenues-lead)/2)
  

# Main plot
#annual_revenues<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(revenues =(revenues) / 10^9,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, revenues)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = revenues, label = project_name),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(8)) +
  expand_limits(y=35)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Annual Revenues (billion dollars)",
    #title = "Annual Revenues by Royalty Project (2016-2023), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(annual_revenues,file="annual_rev_project.png",dpi=500,bg="white",width=10,height=5)

```


```{r annual_real_rev_by_project,fig.align='center',fig.width=12, fig.height=5.5,dpi=500}

# Calculate last value per project
BLCN_labels <- blcn_data %>%
  group_by(project_name) %>%
  mutate(revenues=(revenues)/10^9,
         year=as.numeric(reporting_year)) %>%
  filter(year == max(year)) %>%
  ungroup()%>%
  mutate(n=row_number())%>%arrange(-n)%>%
  mutate(revenues=cumsum(revenues))%>%arrange(n)%>%
  mutate(lead=lead(revenues),lead = replace_na(lead, 0))%>%
  mutate(revenues=revenues-(revenues-lead)/2)
  

# Main plot
#annual_revenues<-
blcn_data %>%
  group_by(project_name) %>%
  mutate(revenues =(revenues) / 10^9,
         year=as.numeric(reporting_year)) %>%
  ungroup() %>%
  ggplot(aes(year, revenues)) +
  geom_col(aes(fill = project_name), size = 0.25, position = "stack", color = "black") +
  geom_text(
    data = BLCN_labels,
    aes(x = year+0.5, y = revenues, label = project_name),
    hjust = 0, size = 3.5, fontface = "bold", show.legend = FALSE
  ) +
  #scale_fill_manual(values = MetBrewer::met.brewer("Signac", 10))+  # or "Signac" "Paquin"
  scale_fill_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
  scale_colour_manual(values = rev(colors_colorblind_10()))+  # or "Signac" "Paquin"
coord_cartesian(clip = "off") +  # allows drawing outside panel
  scale_x_continuous(expand = c(0, 0), breaks = pretty_breaks(10)) +
  scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks(8)) +
  expand_limits(y=35)+
  theme_irpp(base_size = 14) +
  theme(
    #axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5),
    legend.position = "none", # hide legend since we use direct labels
    plot.margin = margin(t = 5, r = 120, b = 5, l = 5)  # right margin expanded (r = 50)
  ) +
  labs(
    x = NULL,
    y = "Annual Revenues (billion dollars)",
    #title = "Annual Revenues by Royalty Project (2016-2023), Beaver Lake Cree Nation Core Territory",
    #caption = "Data via Government of Alberta, calculations and graph by Andrew Leach"
  )

#ggsave(annual_revenues,file="annual_rev_project.png",dpi=500,bg="white",width=10,height=5)

```



## Production comparison

```{r, echo = T,eval=T}
  BLCN_val%>% 
  group_by(year)%>%
  #still monthly production by project (or other) in m3
  summarize(bitumen=sum(bitumen,na.rm = T)/10^6*6.2898,
            value=sum(value)/10^6,
            val_bbl=value/bitumen,
            .groups="drop")%>%
  left_join(
    blcn_data%>%rename(year=reporting_year)%>%
      group_by(year)%>%
      summarize(royalty_prod=sum(prod,na.rm = T)/10^6,
                royalty_val=sum(revenues,na.rm = T)/10^6,
                royalty_val_bbl=royalty_val/royalty_prod,
                .groups="drop"))%>%
  filter(year>=2016)%>%
  mutate(val_diff=royalty_val_bbl-val_bbl,
         mean_val_diff=sum(val_diff*bitumen,na.rm=T)/sum(bitumen),
         prod_diff=bitumen/royalty_prod)%>%
  
  # 
  #   group_by(year)%>%
  # summarize(bitumen=sum(bitumen),days=sum(days),bitumen=bitumen/days)%>%
  # select(Year=year,"Production (mmbbl/d)"=bitumen)%>%
  # mutate(Year=as.character(Year),
  #        Year=ifelse(Year=="2025","2025 (to date)",Year)
  #        )%>%
   kbl(escape = FALSE,table.attr = "style='width:60%;'") %>%
    kable_styling(fixed_thead = T,bootstrap_options = c("hover", "condensed","responsive"),full_width = T)%>%
  #scroll_box(width = "1000px", height = "600px")%>%
  column_spec(1, extra_css = "white-space: nowrap;") %>%
      I() 
```

## Valuation comparison

```{r}
joint<-
  blcn_data %>% #filter(project_name %in% c("Christina Lake (CVE)", "Foster Creek","Christina Lake (MEG)","Cold Lake"))%>%
  rename(year=reporting_year)%>%
  left_join(
    #test<-
    BLCN_val %>%
      #filter(project %in% c("Christina Lake (CVE)", "Foster Creek","Christina Lake (MEG)","Cold Lake"))%>%
               mutate(year=year(date))%>%
              group_by(project,year)%>%
                summarize(prod_st53=sum(bitumen_bbl,na.rm = T),val_st53=sum(value,na.rm = T))%>%
                rename(project_name=project)
              ,by=c("project_name","year"))%>%
  mutate(avg_val_53=val_st53/prod_st53,avg_val_roy=revenues/prod,diff=avg_val_53-avg_val_roy)

joint %>% #filter(year>=2016)%>% group_by(year) %>%
  summarise(val_roy=sum(revenues)/sum(prod),val_53=sum(val_st53)/sum(prod_st53),diff=val_53-val_roy)


```
## Other



```{r} 
## Wells

st37_txt <- read_tsv("../blcn/WellList.txt",col_names = FALSE)%>%
  setNames(c(
  "UWI ID",
  "Loc-Desc",
  "Update-Flag",
  "Well-Name",
  "Field Code",
  "Pool Code",
  "OS-Area- Code",
  "OS-Dep- Code",
  "License-No",
  "License-Status",
  "License-Issue-Date",
  "Licensee-Code",
  "Agent-Code",
  "Operator-Code",
  "Fin-Drl-Date",
  "Well-Total-Depth",
  "Well-Stat-Code",
  "Well-Stat-Date",
  "Fluid_Short_Description",
  "Mode_Short_Description",
  "Type_Short_Description",
  "Structure_Short_Description",
  "Scheme_Type",
  "Scheme_Sub_Type"
))%>%clean_names()





 st37_shp <- st_read("../blcn/ST37_SH/ST37_SH_GCS_NAD83.shp")%>%clean_names()%>%
  st_transform(crs_val)
  
st37_dh <- st_read("../blcn/ST37_BH/ST37_BH_GCS_NAD83.shp")%>%clean_names()%>%
  st_transform(crs_val)


  
st37_blcn<- st_intersection(st37_shp,blcn_buffer)

st37_dh_blcn<- st_intersection(st37_dh,blcn_buffer)

st37_blcn<- st37_blcn %>% group_by(licence,lic_status)%>%mutate(records=n())%>% ungroup()
  


st37_blcn<-st37_blcn%>%left_join(st37_txt%>%select(licence=license_no,fluid_short_description,type_short_description,fin_drl_date)%>%distinct())%>%
  filter(fluid_short_description!="WATER")%>%
  mutate(lic_stat_dat=dmy(lic_stat_dat),
         year=round(year(lic_stat_dat),-1))%>%
  group_by(licence)%>%
  mutate(first=pmin(lic_stat_dat),
         last=pmax(lic_stat_dat))%>%
  ungroup()%>%
  I()
  

non_op<-c("RecExempt","RecCertified","Abandoned","Suspension")

st37_blcn<-st37_blcn%>%
  group_by(licence)%>%
  arrange(lic_stat_dat)%>%
  mutate(last_status=last(lic_status),last_date=last(lic_stat_dat))%>%
  ungroup()%>%
  mutate(op=(!last_status%in%non_op))%>%
  mutate(status=case_when(
    op ~ "Operating at present",
    last_status=="Suspension" ~ "Reclaimed, Abandoned or Suspended since 2008",
    last_status=="Abandoned" ~ "Reclaimed, Abandoned or Suspended since 2008",
    last_status%in%c("RecExempt","RecCertified") ~ "Reclaimed, Abandoned or Suspended since 2008",
    TRUE~"Other"
    ))%>%
  mutate(relevant=(op|last_date>=ymd("2008-01-01")))
    
    
  
    



ggplot(st37_blcn%>%filter(relevant,fluid_short_description %in% c("GAS","CR-BIT","CR-OIL")|type_short_description=="SAGD")) +
  geom_sf(aes(colour=status),size=0.3)+
  geom_sf(data=osb,aes(color = "Oil Sands Project Boundaries"),fill=NA,linewidth=0.5)+
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  geom_sf(data=res_pts,aes(colour="Beaver Lake Cree Nation Reserve"),fill=NA,linewidth=1) +
    #geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
scale_color_manual("",values=colors_tableau10())
  


st37_dh_bl<- st37_dh_blcn %>% mutate(fd_date_n=ymd(fd_date))

```


```{r}
aepa_projects <- read_excel("../blcn/epa-oil-sands-operating-capacity-report-2023-2024.xlsx",sheet="2022 (BY PHASE)") %>% clean_names()

test<-aepa_projects %>% 
  #mutate(project_name=ifelse(grepl("Cold Lake"),"Cold",project_name))%>%
  select(operator,project_name) %>% distinct()
  
  
```



```{r}

access_all<-st_layers("~/data_projects/blcn/Access_GDB_10TM_Forest/bfgen_access_06-11-2024/Access/Access/Access.gdb")

utility_all<-st_layers("~/data_projects/blcn/Access_GDB_10TM_Forest/bfgen_access_06-11-2024/Utility Communication/Utility/Utility.gdb")

roads <- st_read(dsn = "~/data_projects/blcn/Access_GDB_10TM_Forest/bfgen_access_06-11-2024/Access/Access/Access.gdb", layer = "Road",quiet = TRUE)%>%clean_names()%>%
  st_transform(crs_val)

powerlines <- st_read(dsn = "~/data_projects/blcn/Access_GDB_10TM_Forest/bfgen_access_06-11-2024/Utility Communication/Utility/Utility.gdb", layer = "Powerline",quiet = TRUE)%>%clean_names()%>%
  st_transform(crs_val)

powerlines<-st_filter(powerlines, blcn_buffer, .predicate = st_intersects)


rail <- st_read(dsn = "~/data_projects/blcn/Access_GDB_10TM_Forest/bfgen_access_06-11-2024/Access/Access/Access.gdb", layer = "Railway",quiet = TRUE)%>%clean_names()%>%
  st_transform(crs_val)

rail<-st_filter(rail, blcn_terr, .predicate = st_intersects)
roads<-st_filter(roads, blcn_terr, .predicate = st_intersects)

ggplot(rail) +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  geom_sf(colour="firebrick") +
  geom_sf(data=roads%>%filter(hwy_number%in% c("63","881")),aes(color=hwy_number)) +
  geom_sf(data=powerlines, color = "dodgerblue") +
  
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
  geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =NAME),nudge_y=.50)
  
```


```{r}

osb_all <- st_layers("OSProjectsOSIP.gdb")

osb_1998 <- st_read(dsn = "OSProjectsOSIP.gdb", layer = "OSProjectBoundaries1998",quiet = TRUE)%>%clean_names()%>%
  rename(geometry=SHAPE)%>%
  st_transform(crs_val)
osb_1998<-st_filter(osb_1998, blcn_terr, .predicate = st_intersects)


osb_2007 <- st_read(dsn = "OSProjectsOSIP.gdb", layer = "OSProjectBoundaries2007",quiet = TRUE)%>%clean_names()%>%
  rename(geometry=SHAPE)%>%
  st_transform(crs_val)
osb_2007<-st_filter(osb_2007, blcn_terr, .predicate = st_intersects)


osb_2011<-st_read(dsn = "OSProjectsOSIP.gdb", layer = "OSProjectBoundaries2011",quiet = TRUE)%>%clean_names()%>%
  rename(geometry=Shape)%>%
  st_transform(crs_val)
osb_2011<-st_filter(osb_2011, blcn_terr, .predicate = st_intersects)


osb_2015<-st_read(dsn = "OSProjectsOSIP.gdb", layer = "OSProjectBoundaries2015",quiet = TRUE)%>%clean_names()%>%
  rename(geometry=Shape)%>%
  st_transform(crs_val)
osb_2015<-st_filter(osb_2015, blcn_terr, .predicate = st_intersects)

#osb_2015 <- st_read("../blcn/BLCN OSProject Boundaries2015.gpkg",quiet = #
#TRUE)%>%clean_names()

ggplot(osb_2011) +
  geom_sf(aes(), color = "black") +
  #geom_sf(data=in_situ%>%mutate(year=2024),aes(fill = "Current"), color = "white") +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1)


osb<-bind_rows(osb_1998,osb_2007,osb_2011,osb_2015)%>%
  filter(!st_is_empty(geometry)) %>% st_make_valid()%>%
  group_by(osp_no)%>%
  mutate(first_year=as.numeric(min(year)),
         project_name=last(project_name))%>%
  ungroup()%>%
  arrange(operator_name,project_name)%>%
  st_transform(crs_val)%>%
  I()

#sf_use_s2(FALSE)

ggplot(osb) +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  geom_sf(aes(fill = as_factor(first_year)), color = "white") +
  #geom_sf(data=in_situ%>%mutate(year=2024),aes(fill = "Current"), color = "white") +
  
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =             #                                                                     NAME),nudge_y=.50)+
  geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =                                                                                 NAME),nudge_y=10000)+
  scale_fill_viridis("",begin = 0.8,end = 0.1,discrete=T)+
  scale_color_brewer("",palette = "Set1")+
 tombe_theme()+
  theme(
    legend.position = "bottom",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  facet_wrap(~year,nrow = 1)+
  labs(x="",y="",title="Oil Sands Project Boundaries Within BLCN Territory Over Time",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by Andrew Leach")
    
#"~/data_projects/blcn/Great Divide project boundary 2015.gpkg"

```


# Gas



```{r, include=TRUE}
get_pipe_data<-function(pipe_name="Enbridge-Mainline"){
  names<-c("Alliance","Cochin","Enbridge-Mainline","Norman-Wells",
        "Keystone","MNP","Trans-Mountain","TQM","tcpl-mainline","Westcoast","ngtl")
pipe_data<-"No data"
if(pipe_name %in% names)
  {
  file_name<-paste("https://www.cer-rec.gc.ca/open/energy/throughput-capacity/",pipe_name,"-throughput-and-capacity.csv",sep="") 
    local_file<-paste(pipe_name,".csv",sep="")
    download.file(file_name,local_file,mode="wb")
    pipe_data <- read_csv(file = local_file)%>%clean_names()%>%mutate(pipe_name=pipe_name)
  }
pipe_data
}

options(timeout=300)
if (!file.exists("ngtl_data.rdata") || 
    difftime(Sys.time(), file.info("ngtl_data.rdata")$mtime, units = "days") > 10)
  {
  ngtl_data<-get_pipe_data(pipe_name="ngtl")
  #filter(direction_of_flow=="east")
  save(ngtl_data,file = "ngtl_data.rdata")
}
load(file = "ngtl_data.rdata")

```

```{r,echo=T, eval=T,include=TRUE,out.width="95%",dpi=150,fig.align="center",fig.height=7,fig.width=11,cache=FALSE}

ggplot(
  #test<-
  ngtl_data %>%
        #filter(key_point %in% c("OSDA Kirby"))%>%     
        filter(key_point %in% c("OSDA Kirby","OSDA Liege"))%>%
         #filter(date>=ymd("2011-01-01"))%>%
         rename(throughput=throughput_1000_m3_d,
                                                        capacity=capacity_1000_m3_d)) +
  geom_line(aes(date,throughput/1000,color="Throughput"),linewidth=0.85) +
  geom_line(aes(date,capacity/1000,color="Available Capacity"),linewidth=1.25, lty="21") +
  scale_x_date(name=NULL,date_breaks = "1 year", date_labels =  "%b\n%Y",expand=c(0,0)) +
  scale_color_manual("",values=c("black","firebrick"))+
  scale_y_continuous(expand = c(0, 0),
                     
                     sec.axis = sec_axis( trans=~.*0.035314666, name="Throughput or Capacity (Billion Cubic Feet per Day)")) +
  guides(alpha = guide_legend(override.aes = list(fill=viridis(n=3,alpha=1,begin=.8,end=0,option = "E",direction=-1)[1]),order = 10) ,
         fill = guide_legend(order = 1) )+
  expand_limits(y=0)+
  facet_wrap(~key_point,ncol=1)+
  theme_irpp(base_size=14)+
  theme(panel.spacing.y = unit(0.5, "cm"))+
  labs(y="Throughput or Capacity (Millions of Cubic Metres per Day)",x="",
    #title=paste("NGTL System Throughput and Capacity Oil Sands",sep=""),
     #  caption="Source: CER Data for NGTL, graph by Andrew Leach."
    )

```

```{r}
kirby_value<-
  ngtl_data %>%
        filter(key_point %in% c("OSDA Liege"))%>%     
        #filter(key_point %in% c("OSDA Kirby"))%>%
         #filter(date>=ymd("2011-01-01"))%>%
         rename(throughput=throughput_gj_d)%>%
  group_by(month,year)%>%
    summarize(throughput_gj=sum(throughput,na.rm=T),
              date=min(date))%>%
  ungroup()%>%arrange(date)%>%
  select(date,throughput_gj)%>%
  left_join(sproule_hist %>% filter(commodity=="Alberta Spot AECO"))%>%
  left_join(cpi)%>%
  mutate(value_month=price*throughput_gj*0.947817/10^6,
         value_real=value_month/cpi_2025*100,
         disc=2025-year(date)
         )%>% #convert gj to mmbtu and convert to millions
  select(date,throughput_gj,price,value_month,value_real,disc)
  


```
