---
title: "Alberta Oil Sands Royalty Data"
author: Andrew Leach
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      code_folding: hide
      includes:
      after_body: 
      theme: lumen
always_allow_html: yes
editor: source

---
```{r, echo=FALSE,include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,cache = T,fig.align='center',dpi=220,fig.width=16,fig.height=9)
options(scipen = 999)
```

One of the recommendations of the Alberta Royalty Review in 2015-16 was more transparency with respect to oil sands costs, profits, production and royalty payments. The Alberta Government has followed that recommendation and makes [project-level data](https://open.alberta.ca/opendata/alberta-oil-sands-royalty-data1) available on their open data site. This document compiles those data and provides some basic graphs that you might find useful. Where appropriate, I've provided a link to the R code used to process the data.

```{r,cache=F, include=F}
library(janitor)
library(scales)
library(tidyverse)
library(readxl)
library(openxlsx)
library(viridis)
library(ggthemes)
library(kableExtra)
library(sf)

work_theme<-function(){
  theme_tufte(18)+
     theme(
           plot.subtitle = element_text(color="grey10",size=rel(1.25)),
           plot.title = element_text(face="bold"),
           plot.caption = element_text(color="grey50",size=rel(1)),
           legend.title = element_text(color="grey10",size=rel(1)),
           legend.text = element_text(color="grey10",size=rel(1)),
           strip.text = element_text(size=rel(1.2)),
           axis.title = element_text(size=rel(1.2)),
           axis.text = element_text(size=rel(1.2)),
           axis.text.x = element_text(size=rel(.8)),
           axis.ticks = element_blank(),
           panel.spacing = unit(.75,"lines"),
           legend.position = "bottom",
           plot.margin = margin(t = .5, r = .5, b = .5, l = .5,unit= "cm"),
           #axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)
           NULL
           )+
  NULL
}

colors_tableau10<-function () 
{
    return(c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", 
        "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17BECF"))
}

```

<!-- ## Download and Combine the Annual Data for 2016-2022 -->

<!-- The first step is to load and process the data. The code button provides a window into how I've done this. In this case, I've manually combined the data in my own spreadsheet so I can see any changes and also look up individual project information easily. -->


```{r load data,cache=T,include=F}
#load plant data
make_os_data<-function(){

os_data_2023 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2023 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2022 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2022 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
os_data_2021 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2021 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2020 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2020 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2019 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2019 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  os_data_2018 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2018 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2017 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2017 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
os_data_2016 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2016 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

#stack all the years together

os_data<-bind_rows(os_data_2023,os_data_2022,os_data_2021,os_data_2020,os_data_2019,os_data_2018,os_data_2017,os_data_2016)

#clean the names

os_data<-os_data %>% clean_names()%>%
  mutate(op_profit=gross_revenue-operating_costs-other_costs+other_net_proceeds)


# data are reported for an individual project twice if it passes payout in a given year

# see here for a test
#os_data %>% group_by(project,reporting_year) %>% select(project,project_name,reporting_year,payout_status) %>% mutate(n_obs=n())%>% filter(n_obs>1) %>% arrange(-n_obs)

# combine the data and include two things: a payout year indicator and a net and gross rev royalty number along with a total royalty paid
os_data <- os_data %>% group_by(project,reporting_year,project_name,operator_name)%>%
  summarize(
    project_revenue=sum(project_revenue),
    gross_revenue=sum(gross_revenue),
    op_profit=sum(op_profit),
    cleaned_crude_bitumen_at_rcp_barrels=sum(cleaned_crude_bitumen_at_rcp_barrels),
    operating_costs=sum(operating_costs),
    capital_costs=sum(capital_costs),
    return_allowance=sum(return_allowance),
    other_costs=sum(other_costs),
    other_net_proceeds=sum(other_net_proceeds),
    net_revenue=sum(net_revenue),
    royalty_type=last(royalty_type),
    royalty_type_start=first(royalty_type),
    payout_start=first(payout_status),
    payout_status=last(payout_status),
    net_rev_royalty=sum(royalty_payable*(royalty_type=="NET")),
    gross_rev_royalty=sum(royalty_payable*(royalty_type=="GROSS")),
    royalty_payable=sum(royalty_payable),
    royalty_rate=last(royalty_rate_percent),
    first_royalty_rate=first(royalty_rate_percent),
    royalty_avg=royalty_payable/gross_revenue,
    payout_year=(n()>1),
    unrecovered_balance_net_loss_at_eop=last(unrecovered_balance_net_loss_at_eop))%>%
    mutate(
      last_prod=last(cleaned_crude_bitumen_at_rcp_barrels)/365,
      label=paste(project_name,"\n(",formatC(round(last_prod,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""))%>%
  ungroup()%>%
  group_by(project)%>%
  mutate(project_payout=(last(payout_status)=="POST"))

#fix facilities
os_data <- os_data %>% mutate(
  royalty_bbl=royalty_payable/cleaned_crude_bitumen_at_rcp_barrels,
  op_costs_bbl=operating_costs/cleaned_crude_bitumen_at_rcp_barrels,
  gross_revenue_bbl=gross_revenue/cleaned_crude_bitumen_at_rcp_barrels,
  cap_costs_bbl=capital_costs/cleaned_crude_bitumen_at_rcp_barrels,
  project_name=gsub(" Project","",project_name),
  project_name=gsub("Christina Lake Regional","Christina Lake (MEG)",project_name),
  project_name=gsub("Christina Lake Thermal","Christina Lake (CVE)",project_name),
  project_name=gsub("MacKay River Commercial","PetroChina",project_name),
  project_name=gsub("MacKay River","MacKay River (Suncor)",project_name),
  project_name=gsub("PetroChina","MacKay River (PetroChina)",project_name),
  project_name=gsub("Harvest BlackGold","BlackGold",project_name),
  
  project_name=gsub(" Thermal","",project_name),
  project_name=gsub(" Mine","",project_name),
  project_name=gsub(" Oil Sands","",project_name),
  project_name=gsub(" EOR","",project_name),
  project_name=gsub(" Commercial","",project_name),
  project_name=gsub(" SAGD","",project_name),
  project_name=gsub(" Demonstration","",project_name),
  project_name=gsub(" In-Situ","",project_name),
  project_name=as.factor(project_name),
  op_profit_net=op_profit-royalty_payable,
  op_profit_bbl=op_profit/cleaned_crude_bitumen_at_rcp_barrels,
  op_profit_net_bbl=op_profit_net/cleaned_crude_bitumen_at_rcp_barrels
  )

}

os_data<-make_os_data()

    
mines<-c("Muskeg River","Fort Hills","Kearl","Horizon","Muskeg River","Jackpine","Syncrude","Suncor")

SAGD<-c("Hangingstone","Leismer","Blackrod","Mackay River","Kirby","Christina Lake","Foster Creek",
        "Long Lake","Great Divide","Surmont","Jackfish","Sunrise","Orion","Firebag","MacKay River",
        "West Ells","SAGD","Tucker","Narrows Lake")

BLCN<-c("Blackrod",
        "BlackGold",
        "Christina Lake (CVE)",
        "Christina Lake (MEG)",
        "Cold Lake",
        "Foster Creek",
        "Jackfish",
        "Kirby North",
        "Kirby South",
        "Leismer",
        "Wolf Lake Crown Agreement",
        "Burnt Lake",
        "Primrose",
        "Narrows Lake",
        "Orion",
        "Pike 1",
        "Surmont",
        "Tucker"
        )

os_data<-os_data %>% mutate(blcn=(project_name %in% BLCN),
                            )%>%
  group_by(project_name)%>%
  mutate(min_bbls=min(cleaned_crude_bitumen_at_rcp_barrels)/365,
         max_bbls=max(cleaned_crude_bitumen_at_rcp_barrels)/365)%>%
  ungroup()

blcn_data<-
      os_data %>% filter(blcn)%>%
         mutate(project_name=as.factor(project_name),
                project_name=fct_reorder(project_name,-last_prod)
                )
write.csv(file = "projects.csv",blcn_data%>%select(project_name)%>%distinct())

blcn_data<-blcn_data%>%
        mutate(project_name=fct_other(project_name,keep=head(levels(blcn_data$project_name),5),other_level = "All Other Projects Combined"))%>%
        group_by(project_name,reporting_year)%>%
        summarize(prod=sum(cleaned_crude_bitumen_at_rcp_barrels,na.rm = T),
                  royalties=sum(royalty_payable,na.rm = T),
                  revenues=sum(gross_revenue,na.rm=T)
                    )%>%
  ungroup()
  

save(blcn_data,file = "os_data_blcn.rdata")

total<-blcn_data %>% 
         summarize(royalties=sum(royalties,na.rm=T)/10^9,
                   rev=sum(revenues,na.rm=T)/10^9,
                   prod=sum(prod,na.rm=T)/10^9)


#os_data<-os_data%>%select(project_name,mine,big_project,reporting_year,production=cleaned_crude_bitumen_at_rcp_barrels,gross_revenue,capital_costs,operating_costs,other_net_proceeds,other_costs,royalty_payable,payout_status)

#save(os_data,file = "os_data.rdata")

```

## Projects Table
```{r prj_table,fig.align='center',fig.width=16, fig.height=8,dpi=220}

  # blcn_data%>%
  # mutate(prod=prod/(365+leap_year(reporting_year)*1))%>%
  # kbl(escape = FALSE,table.attr = "style='width:80%;'",digits=2,align=rep('c', 2)) %>%
  # kable_styling(fixed_thead = T,bootstrap_options = c("hover", "condensed","responsive"),full_width = T)%>%
  # add_header_above(header = c("Western Canadian Select (WCS) prices in the Canada Net-Zero scenario of the Canada Energy Regulator's Canada's Energy Future (2023) report"=2))%>%
  # I() 
  #     ggplot(aes(factor(reporting_year),prod/10^3))+
  # geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  # #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  # scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  # #coord_flip()+
  # scale_fill_brewer()+
  # guides(fill="none",color="none")+
  # facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  # #scale_x_reverse()+
  # work_theme()+
  # theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
  #       )+
  # labs(x=NULL,y="Bitumen Production (1000 barrels per day)",
  #      title="Annual Bitumen Production by Project, Beaver Lake Cree Nation Core Territory",
  #      subtitle=paste("Total production since 2016 of ",round(total$prod,2)," billion barrels and 2023 average production of ",sum(blcn_data%>%filter(reporting_year==2023)%>%select(prod)/365/10^6)%>%round(2)," million barrels per day",sep=""),
  #      caption="Data via Government of Alberta, graph by @andrew_leach")
  #      
  # 

```



## Production

```{r in situ prod,fig.align='center',fig.width=16, fig.height=8,dpi=220}

  blcn_data%>%
  mutate(prod=prod/(365+leap_year(reporting_year)*1))%>%
      ggplot(aes(factor(reporting_year),prod/10^3))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Bitumen Production (1000 barrels per day)",
       title="Annual Bitumen Production by Project, Beaver Lake Cree Nation Core Territory",
       subtitle=paste("Total production since 2016 of ",round(total$prod,2)," billion barrels and 2023 average production of ",sum(blcn_data%>%filter(reporting_year==2023)%>%select(prod)/365/10^6)%>%round(2)," million barrels per day",sep=""),
       caption="Data via Government of Alberta, graph by @andrew_leach")
       
  

```


```{r cum_prod,fig.align='center',fig.width=16, fig.height=8,dpi=220}

        blcn_data%>%
        group_by(project_name)%>%
        mutate(prod=cumsum(prod))%>%
        ungroup()%>%
  filter(reporting_year==2023)%>%
        ggplot(aes(factor(reporting_year),prod/10^6))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  #expand_limits(y=30)+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  #theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)     )+
  theme(axis.text.x = element_blank())+
  labs(x=NULL,y="Production (million barrels)",
       title="Total Bitumen Production (2016-2023), Beaver Lake Cree Nation Core Territory",
       subtitle=paste("Total production from all projects in BLCN Territory since 2016 of ",round(total$prod,2)," billion barrels",sep=""),
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by Andrew Leach")

```



## Project Revenue

```{r in situ prod,fig.align='center',fig.width=16, fig.height=8,dpi=220}

        blcn_data%>%
       ggplot(aes(factor(reporting_year),revenues/10^9))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  expand_limits(y=8)+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Project Revenue (billions of dollars per year)",
       title="Annual Gross Bitumen Revenue, Larger Oil Sands In Situ Projects",
       subtitle=paste("Total project revenue since 2016 of CA$",round(total$rev,2)," billion, or CA$",round(total$rev/total$prod,2)," per barrel on average.",sep=""),
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

```{r cum_rev,fig.align='center',fig.width=16, fig.height=8,dpi=220}

        blcn_data%>%
        group_by(project_name)%>%
        mutate(revenues=cumsum(revenues))%>%
  filter(reporting_year==2023)%>%
        ggplot(aes(factor(reporting_year),revenues/10^9))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  expand_limits(y=30)+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  #theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)     )+
  theme(axis.text.x = element_blank())+
  labs(x=NULL,y="Project Revenue (billions of dollars, 2016-2023)",
       title="Total Bitumen Revenue (2016-2023) by Project, Beaver Lake Cree Nation Core Territory",
       subtitle=paste("Total revenue from all projects in BLCN Territory since 2016 of CA$",round(total$rev,2)," billion",sep=""),
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by Andrew Leach")

```




## Royalties

```{r cum_roy,fig.align='center',fig.width=16, fig.height=8,dpi=220}

        blcn_data%>%
        group_by(project_name)%>%
        mutate(royalties=cumsum(royalties))%>%
  filter(reporting_year==2023)%>%
        ggplot(aes(factor(reporting_year),royalties/10^9))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  #expand_limits(y=30)+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  #theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)     )+
  theme(axis.text.x = element_blank())+
  labs(x=NULL,y="Royalty Payments (CA$ billions, 2016-2023)",
       title="Total Royalties Paid (2016-23) by Project, Beaver Lake Cree Nation Core Territory",
       subtitle=paste("Projects in BLCN Territory have paid CA$",round(total$royalties,2)," billion in royalties since 2016 (CA$",sum(blcn_data%>%filter(reporting_year==2023)%>%select(royalties)/10^9)%>%round(2)," billion paid in 2023)" ,sep=""),
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by Andrew Leach")

```




```{r pipes,fig.align='center',fig.width=16, fig.height=8,dpi=220}

#gd_pts <- st_read("../blcn/Pipelines_SHP/Pipelines_GCS_NAD83.shp")%>%clean_names()

gd_pts <- st_read("../blcn/Pipelines_SHP/Pipelines_NAD83_10TM_AEPForest.shp")%>%clean_names()%>%
  st_transform(crs = 4326)

```

```{r reserves,fig.align='center',fig.width=16, fig.height=8,dpi=220}

#gd_pts <- st_read("../blcn/Pipelines_SHP/Pipelines_GCS_NAD83.shp")%>%clean_names()

res_pts <- st_read("../blcn/ILC/Indigenous_Lands_of_Canada.shp")%>%clean_names()%>%
  st_transform(crs = 4326)%>%
  filter(grepl("BEAVER LAKE 131",admin_area_n))

```



```{r proj_map,fig.align='center',fig.width=16, fig.height=8,dpi=220}
gd_clean<-gd_pts %>%
  filter(substance1 %in% c("Natural Gas","Crude Oil","Hydrocarbon Diluents"))

gd_crude<-gd_pts %>%
  filter(substance1 %in% c("Crude Oil"))


larp_pop<- st_read("../blcn/LARP/LA_RAC_Recommended_Population_Centers.shp")%>%
  st_transform(crs=st_crs(gd_pts))
larp_pop<- st_read("../blcn/LARP/LA_RAC_Recommended_Population_Centers.shp")%>%
  st_transform(crs=st_crs(gd_pts))

larp_bd <-st_read("../blcn/LARP/LUF Integrated Regional Plan Boundaries.shp")%>%
  st_transform(crs=st_crs(gd_pts))

l_ath_bd <-  st_intersection(gd_clean,larp_bd %>% 
  filter(LUF_CODE=="02"))


blcn_terr <-st_read("../blcn/blcn_poly/Beaver Lake Cree Traditional Territory.shp")%>%
  st_transform(crs=st_crs(gd_pts))

blcn_buffer<-st_buffer(blcn_terr, units::as_units(200, 'kilometer'))
oil_blcn<-st_intersection(gd_crude,blcn_buffer)

#oil_blcn<-gd_crude#%>%filter(from_fac=="Pipeline",as.numeric(out_diamet)>100)

ggplot(l_ath_bd)+
  geom_sf(data=oil_blcn,aes(colour="Crude Oil Pipelines"),linewidth=2,fill=NA) +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=5) +
  geom_sf(data=res_pts,aes(colour="Beaver Lake #131 Reserve"),fill="grey80",linewidth=1) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,11,12,15)))+
  geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,11,12,15)), aes(label = NAME),nudge_y=.20)+
  scale_colour_brewer("",palette = "Set1")+
  theme_minimal()
  #scale_fill_manual("")+
 theme_void()+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  labs(x="",y="",
      title="Crude Oil and Gas Pipeline Infrastructure Within BLCN Territory",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    

```




```{r proj_map,fig.align='center',fig.width=16, fig.height=8,dpi=220}

osb_1980 <- st_read("../blcn/BLCN OSProject Boundaries1980s.gpkg")%>%clean_names()

osb_1998 <- st_read("../blcn/BLCN OSProject Boundaries1998.gpkg")%>%clean_names()

osb_2007 <- st_read("../blcn/BLCN OSProject Boundaries2007.gpkg")%>%clean_names()

osb_2011 <- st_read("../blcn/BLCN OSProject Boundaries2011.gpkg")%>%clean_names()

osb_2015 <- st_read("../blcn/BLCN OSProject Boundaries2015.gpkg")%>%clean_names()

osb<-bind_rows(osb_1980,osb_1998,osb_2007,osb_2015)%>%clean_names()%>%
  group_by(osp_no)%>%
  mutate(first_year=as.numeric(min(year)),
         project_name=last(project_name))%>%
  ungroup()%>%
  arrange(operator_name,project_name)

#write.csv(file = "projects_osip.csv",osb%>%st_drop_geometry()%>% #select(project_name,operator_name)%>%distinct())



ggplot(osb) +
  geom_sf(aes(fill = as_factor(first_year)), color = "white") +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =             #                                                                     NAME),nudge_y=.50)+
  geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =                                                                                 NAME),nudge_y=10000)+
  scale_fill_viridis("",begin = 0.8,end = 0.1,discrete=T)+
  scale_color_brewer("",palette = "Set1")+
 tombe_theme()+
  theme(
    legend.position = "bottom",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  facet_wrap(~year,nrow = 1)+
  labs(x="",y="",title="Oil Sands Project Boundaries Within BLCN Territory Over Time",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    
#"~/data_projects/blcn/Great Divide project boundary 2015.gpkg"

```


```{r}
aepa_projects <- read_excel("../blcn/epa-oil-sands-operating-capacity-report-2023-2024.xlsx",sheet="2022 (BY PHASE)") %>% clean_names()

test<-aepa_projects %>% 
  #mutate(project_name=ifelse(grepl("Cold Lake"),"Cold",project_name))%>%
  select(operator,project_name) %>% distinct()
  
  
```




```{r,fig.align='center',fig.width=16, fig.height=8,dpi=220}
all_projects <- st_read("../blcn/OSProjects2015.gdb")%>%clean_names()%>%
  st_transform(crs=st_crs(gd_pts))%>% st_make_valid()%>%
  filter(!project_name%in%c("Caribou","Grouse","Walleye","Birchwood",
                            "Chard","Winefred Lake","West Kirby",
                            "Mariana - Thornbury","South Leismer","Sage"))%>% #cancelled
  filter(!project_name%in%c("Whitesands","Foster Creek North","Jackfish East"))%>%
  mutate(project_name=gsub("Kirby Expansion Project","Kirby",project_name))%>%
  mutate(project_name=gsub("Tucker Lake","Tucker",project_name))%>%
  mutate(project_name=gsub("Cold Lake Operations","Cold Lake (Imperial Oil)",project_name))%>%
  filter(!(operator_name=="MEG Energy Corp."&project_name=="Surmont"))%>%
  filter(!(operator_name=="MEG Energy Corp."&project_name=="May River")) #cancelled


blcn_buffer_50<-st_buffer(blcn_terr, units::as_units(50, 'kilometer'))%>%st_make_valid()%>%
  st_transform(crs=st_crs(gd_pts))

os_blcn_strict<-st_intersection(all_projects,blcn_terr)

os_blcn<-st_intersection(all_projects,blcn_buffer_50)%>%
  filter(project_name %in% os_blcn_strict$project_name)%>%
  #mutate(label=paste(project_name,"\n(",operator_name,")",sep=""))
  mutate(label=project_name)

library(ggrepel)
#test<-"tar"

ggplot(blcn_buffer_50) +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
#  geom_sf(data=all_projects %>% filter(grepl(test, project_name)),aes(fill = as_factor(project_name)), color = "white") +
 # geom_sf_label(data = all_projects %>% filter(grepl(test, project_name)), aes(label =project_name),size=1)+
geom_sf(data=os_blcn,aes(),fill="grey40", color = "white") +
  ggrepel::geom_label_repel(
    data = os_blcn,
    aes(label = label, geometry = Shape),size=3.5,
    stat = "sf_coordinates",
    min.segment.length = 0
  )+
  ggrepel::geom_label_repel(
    data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)),
    aes(label = NAME,geometry=geometry),size=3.5, color="red",
    stat = "sf_coordinates",
    min.segment.length = 10,
    nudge_y=-.05
  )+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =                        #                                                         NAME),nudge_y=.10,color="blue")+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =             #                                                                     NAME),nudge_y=.50)+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =                        #                                                         NAME),nudge_y=10000)+
  scale_fill_viridis("",begin = 0.8,end = 0.1,discrete=T,guide=FALSE)+
  scale_color_brewer("",palette = "Set1")+
 tombe_theme()+
  theme(
    legend.position = "bottom",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  #facet_wrap(~year,nrow = 1)+
  labs(x="",y="",title="Oil Sands Projects Within BLCN Territory",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    
#"~/data_projects/blcn/Great Divide project boundary 2015.gpkg"

```



```{r proj_map,fig.align='center',fig.width=16, fig.height=8,dpi=220}

gd_pts <- st_read("../blcn/Pipelines_SHP/Pipelines_GCS_NAD83.shp")%>%clean_names()
#st_crs(gd_pts)

gd_clean<-gd_pts %>%
  filter(substance1 %in% c("Natural Gas","Crude Oil","Hydrocarbon Diluents"))

larp_pop<- st_read("../blcn/LARP/LA_RAC_Recommended_Population_Centers.shp")%>%
  st_transform(crs=st_crs(gd_pts))
larp_bd <-st_read("../blcn/LARP/LUF Integrated Regional Plan Boundaries.shp")%>%
  st_transform(crs=st_crs(gd_pts))%>% st_make_valid()

l_ath_bd <-  st_intersection(gd_clean,larp_bd %>% 
  filter(LUF_CODE=="02"))

ggplot(l_ath_bd)+
  geom_sf(aes(colour=substance1),fill="white") +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,15)))+
  geom_sf_text(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,15)), aes(label = NAME))+
  scale_colour_brewer("",palette = "Set1")+
  scale_fill_manual("")+
 theme_void()+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  labs(x="",y="",
      title="Pipeline Infrastructure Within BLCN Territory Over Time",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    

```



```


