---
title: "Leach BLCN Report Data And Mapping Compilation"
author: Andrew Leach
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      code_folding: hide
      includes:
      after_body: 
      theme: lumen
always_allow_html: yes
editor: source

---
```{r, echo=FALSE,include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,cache = T,fig.align='center',dpi=220,fig.width=16,fig.height=9)
options(scipen = 999)
```



```{r,cache=F, include=F}
library(janitor)
library(scales)
library(tidyverse)
library(readxl)
library(openxlsx)
library(viridis)
library(ggthemes)
library(kableExtra)
library(sf)
library(cowplot)
library(ggrepel)

work_theme<-function(){
  theme_tufte(18)+
     theme(
           plot.subtitle = element_text(color="grey10",size=rel(1.25)),
           plot.title = element_text(face="bold"),
           plot.caption = element_text(color="grey50",size=rel(1)),
           legend.title = element_text(color="grey10",size=rel(1)),
           legend.text = element_text(color="grey10",size=rel(1)),
           strip.text = element_text(size=rel(1.2)),
           axis.title = element_text(size=rel(1.2)),
           axis.text = element_text(size=rel(1.2)),
           axis.text.x = element_text(size=rel(.8)),
           axis.ticks = element_blank(),
           panel.spacing = unit(.75,"lines"),
           legend.position = "bottom",
           plot.margin = margin(t = .5, r = .5, b = .5, l = .5,unit= "cm"),
           #axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)
           NULL
           )+
  NULL
}

colors_tableau10<-function () 
{
    return(c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", 
        "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17BECF"))
}

```

All data on royalties for 2016-2023 are posted by the Government of Alberta [here](https://open.alberta.ca/opendata/alberta-oil-sands-royalty-data1). I have compiled each of the annual files into a single spreadsheet for use for my report. The code below processes the royalty data for use in my expert report for *Beaver Lake Cree Nation et al v. His Majesty the King in Right of the Province of Alberta, et al.*, Court File No.: 0803 06718, Alberta Court of King's Bench. 

<!-- ## Download and Combine the Annual Data for 2016-2022 -->

```{r load data,cache=T,include=F}
#load plant data
make_os_data<-function(){

os_data_2023 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2023 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2022 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2022 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
os_data_2021 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2021 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2020 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2020 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2019 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2019 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  os_data_2018 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2018 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2017 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2017 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
os_data_2016 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2016 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

#stack all the years together

os_data<-bind_rows(os_data_2023,os_data_2022,os_data_2021,os_data_2020,os_data_2019,os_data_2018,os_data_2017,os_data_2016)

#clean the names

os_data<-os_data %>% clean_names()%>%
  mutate(op_profit=gross_revenue-operating_costs-other_costs+other_net_proceeds)

# data are reported for an individual project twice if it passes payout in a given year
# combine the data and include two things: a payout year indicator and a net and gross rev royalty number along with a total royalty paid
os_data <- os_data %>% group_by(project,reporting_year,project_name,operator_name)%>%
  summarize(
    project_revenue=sum(project_revenue),
    gross_revenue=sum(gross_revenue),
    op_profit=sum(op_profit),
    cleaned_crude_bitumen_at_rcp_barrels=sum(cleaned_crude_bitumen_at_rcp_barrels),
    operating_costs=sum(operating_costs),
    capital_costs=sum(capital_costs),
    return_allowance=sum(return_allowance),
    other_costs=sum(other_costs),
    other_net_proceeds=sum(other_net_proceeds),
    net_revenue=sum(net_revenue),
    royalty_type=last(royalty_type),
    royalty_type_start=first(royalty_type),
    payout_start=first(payout_status),
    payout_status=last(payout_status),
    net_rev_royalty=sum(royalty_payable*(royalty_type=="NET")),
    gross_rev_royalty=sum(royalty_payable*(royalty_type=="GROSS")),
    royalty_payable=sum(royalty_payable),
    royalty_rate=last(royalty_rate_percent),
    first_royalty_rate=first(royalty_rate_percent),
    royalty_avg=royalty_payable/gross_revenue,
    payout_year=(n()>1),
    unrecovered_balance_net_loss_at_eop=last(unrecovered_balance_net_loss_at_eop))%>%
    mutate(
      last_prod=last(cleaned_crude_bitumen_at_rcp_barrels)/365,
      label=paste(project_name,"\n(",formatC(round(last_prod,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""))%>%
  ungroup()%>%
  group_by(project)%>%
  mutate(project_payout=(last(payout_status)=="POST"))

#fix facility names
os_data <- os_data %>% mutate(
  royalty_bbl=royalty_payable/cleaned_crude_bitumen_at_rcp_barrels,
  op_costs_bbl=operating_costs/cleaned_crude_bitumen_at_rcp_barrels,
  gross_revenue_bbl=gross_revenue/cleaned_crude_bitumen_at_rcp_barrels,
  cap_costs_bbl=capital_costs/cleaned_crude_bitumen_at_rcp_barrels,
  project_name=gsub(" Project","",project_name),
  project_name=gsub("Christina Lake Regional","Christina Lake (MEG)",project_name),
  project_name=gsub("Christina Lake Thermal","Christina Lake (CVE)",project_name),
  project_name=gsub("MacKay River Commercial","PetroChina",project_name),
  project_name=gsub("MacKay River","MacKay River (Suncor)",project_name),
  project_name=gsub("PetroChina","MacKay River (PetroChina)",project_name),
  project_name=gsub("Harvest BlackGold","BlackGold",project_name),
  
  project_name=gsub(" Thermal","",project_name),
  project_name=gsub(" Mine","",project_name),
  project_name=gsub(" Oil Sands","",project_name),
  project_name=gsub(" EOR","",project_name),
  project_name=gsub(" Commercial","",project_name),
  project_name=gsub(" SAGD","",project_name),
  project_name=gsub(" Demonstration","",project_name),
  project_name=gsub(" In-Situ","",project_name),
  project_name=as.factor(project_name),
  op_profit_net=op_profit-royalty_payable,
  op_profit_bbl=op_profit/cleaned_crude_bitumen_at_rcp_barrels,
  op_profit_net_bbl=op_profit_net/cleaned_crude_bitumen_at_rcp_barrels
  )

}

os_data<-make_os_data()


BLCN<-c("Blackrod",
        "BlackGold",
        "Christina Lake (CVE)",
        "Christina Lake (MEG)",
        "Cold Lake",
        "Foster Creek",
        "Jackfish",
        "Kirby North",
        "Kirby South",
        "Leismer",
        "Wolf Lake Crown Agreement",
        "Burnt Lake",
        "Primrose",
        "Narrows Lake",
        "Orion",
        "Pike 1",
        "Surmont",
        "Tucker"
        )

os_data<-os_data %>% mutate(blcn=(project_name %in% BLCN),
                            )%>%
  group_by(project_name)%>%
  mutate(min_bbls=min(cleaned_crude_bitumen_at_rcp_barrels)/365,
         max_bbls=max(cleaned_crude_bitumen_at_rcp_barrels)/365)%>%
  ungroup()

blcn_data<-
      os_data %>% filter(blcn)%>%
         mutate(project_name=as.factor(project_name),
                project_name=fct_reorder(project_name,-last_prod)
                )
#make spreadsheet file for report writing
write.csv(file = "projects.csv",blcn_data%>%select(project,project_name,operator_name)%>%
            group_by(project)%>%
            summarize(project_name=last(str_to_title(project_name)),operator_name=last(str_to_title(operator_name))))

#combined projects keeping only the largest 5 and capturing other projects in one category

blcn_data<-blcn_data%>%
        mutate(project_name=fct_other(project_name,keep=head(levels(blcn_data$project_name),5),other_level = "All Other Projects Combined"))%>%
        group_by(project_name,reporting_year)%>%
        summarize(prod=sum(cleaned_crude_bitumen_at_rcp_barrels,na.rm = T),
                  royalties=sum(royalty_payable,na.rm = T),
                  revenues=sum(gross_revenue,na.rm=T)
                    )%>%
  ungroup()
  
#save the data file
save(blcn_data,file = "os_data_blcn.rdata")

#summarize the total for these projects
total<-blcn_data %>% 
         summarize(royalties=sum(royalties,na.rm=T)/10^9,
                   rev=sum(revenues,na.rm=T)/10^9,
                   prod=sum(prod,na.rm=T)/10^9)
```

## Production

```{r in situ prod,fig.align='center',fig.width=12, fig.height=5,dpi=220}
 

 blcn_data%>%
  mutate(prod=prod/(365+leap_year(reporting_year)*1))%>%
      ggplot(aes(factor(reporting_year),prod/10^3))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  theme_ps_grid()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Bitumen Production (1000 barrels per day)",
       title="Annual Bitumen Production by Project, Beaver Lake Cree Nation Core Territory",
       subtitle=paste("Total production since 2016 of ",round(total$prod,2)," billion barrels and 2023 average production of ",sum(blcn_data%>%filter(reporting_year==2023)%>%select(prod)/365/10^6)%>%round(2)," million barrels per day",sep=""),
       caption="Government of Alberta royalty data, graph by @andrew_leach")
       
  

```

## Project Revenue

```{r in situ prod,fig.align='center',fig.width=12, fig.height=5,dpi=220}

        blcn_data%>%
       ggplot(aes(factor(reporting_year),revenues/10^9))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  scale_fill_brewer()+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  expand_limits(y=8)+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  theme_ps_grid()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Project Revenue (billions of dollars per year)",
       title="Annual Gross Bitumen Revenue, Beaver Lake Cree Nation Core Territory",
       subtitle=paste("Total project revenue since 2016 of CA$",round(total$rev,2)," billion, or CA$",round(total$rev/total$prod,2)," per barrel on average.",sep=""),
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

## Royalties

```{r cum_roy,fig.align='center',fig.width=12, fig.height=5,dpi=220}

        blcn_data%>%
        #group_by(project_name)%>%
        #mutate(royalties=cumsum(royalties))%>%
  #filter(reporting_year==2023)%>%
        ggplot(aes(factor(reporting_year),royalties/10^9))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  theme_ps_grid()+
  scale_y_continuous(breaks=pretty_breaks(n=5))+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)     )+
  #theme(axis.text.x = element_blank())+
  labs(x=NULL,y="Royalty Payments (CA$ billions, 2016-2023)",
       title="Annual Royalties Paid (2016-23) by Project, Beaver Lake Cree Nation Core Territory",
       subtitle=paste("Projects in BLCN Territory have paid CA$",round(total$royalties,2)," billion in royalties since 2016 (CA$",sum(blcn_data%>%filter(reporting_year==2023)%>%select(royalties)/10^9)%>%round(2)," billion paid in 2023)" ,sep=""),
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by Andrew Leach")

```

# Territory Shape Files

```{r reserves,fig.align='center',fig.width=16, fig.height=8,dpi=220}

#gd_pts <- st_read("../blcn/Pipelines_SHP/Pipelines_GCS_NAD83.shp")%>%clean_names()

res_pts <- st_read("../blcn/ILC/Indigenous_Lands_of_Canada.shp")%>%clean_names()%>%
  st_transform(crs = 3347)%>%
  filter(grepl("BEAVER LAKE 131",admin_area_n))


larp_pop<- st_read("../blcn/LARP/LA_RAC_Recommended_Population_Centers.shp")%>%
  st_transform(3347)

blcn_terr <-st_read("../blcn/blcn_poly/Beaver Lake Cree Traditional Territory.shp")%>%
  st_transform(3347)

# Buffer in meters (20 km = 20,000 meters)
blcn_buffer <- st_buffer(blcn_terr, dist = 20000)

# Buffer in meters (500 km = 500,000 meters)
blcn_buffer_500 <- st_buffer(blcn_terr, dist = 500000)


```

```{r clawr}
library(httr)
url <- "https://geospatial.alberta.ca/titan/rest/services/boundary/military_base_and_training_ground/MapServer/0/query"

# Set up the query parameters
params <- list(
  where = "1=1",
  outFields = "*",
  f = "geojson"
)

# Make the GET request
response <- GET(url, query = params)

# Check if the request was successful
if (status_code(response) == 200) {
  # Parse the GeoJSON content
  geojson_content <- content(response, as = "text", encoding = "UTF-8")
  clawr_data <- st_read(geojson_content,quiet = TRUE)%>%
  st_transform(3347)
} else {
  stop("Failed to retrieve data from the REST service.")
}


```


# Pipelines in BLCN Core Territory

```{r pipes,fig.align='center',fig.width=16, fig.height=8,dpi=220}

if(!file.exists("Pipeline_Inst.zip")){
  download.file("https://www.aer.ca/data/pipeline/Pipeline_Installations_SHP.zip","Pipeline_Inst.zip",mode="wb")
unzip("Pipeline_Inst.zip")
}

if(!file.exists("Pipelines_SHP.zip")){
download.file("https://www.aer.ca/data/pipeline/Pipelines_SHP.zip","Pipelines_SHP.zip",mode="wb")
unzip("Pipelines_SHP.zip")
}

pipes <- st_read("Pipelines_SHP/Pipelines_NAD83_10TM_AEPForest.shp")%>%clean_names()%>%
  filter(substance1 %in% c("Natural Gas","Crude Oil","Hydrocarbon Diluents"))%>%
    st_transform(crs = 3347)

inst_pts <- st_read("Pipeline_Installations_SHP/Pipeline_Installations_NAD83_10TM_AEPForest.shp",quiet=TRUE)%>%clean_names()%>%
  st_transform(crs = 3347)


inst_pts<-st_filter(inst_pts, blcn_buffer, .predicate = st_intersects)
  

#blcn_buffer<-st_buffer(blcn_terr, units::as_units(200, 'kilometer'))
oil_blcn<-st_filter(pipes, blcn_buffer, .predicate = st_intersects)%>%
  filter(substance1=="Crude Oil")

oil_l_ath_bd<-st_filter(pipes, l_ath_bd, .predicate = st_intersects)%>%
  filter(substance1=="Crude Oil")

oil_pipes<-pipes%>%
  filter(substance1=="Crude Oil")%>%
  filter(seg_status=="Operating")


#oil_blcn<-gd_crude#%>%filter(from_fac=="Pipeline",as.numeric(out_diamet)>100)
```

```{r pipe_graph,fig.align='center',fig.width=12, fig.height=7,dpi=220}

# Create a bounding box in lon/lat (EPSG:4326)
bbox_deg <- st_as_sfc(st_bbox(c(
  xmin = -113.5,
  xmax = -108,
  ymin = 54.5,
  ymax = 56.75
), crs = 4326))

# Step 2: Transform to CRS 3347 (in metres)
bbox_m <- st_transform(bbox_deg, crs = 3347)

# Step 3: Extract the numeric limits in metres
bbox_coords <- st_bbox(bbox_m)


ggplot(l_ath_bd)+
  geom_sf(data=oil_pipes,aes(color="Crude Oil Pipelines"),linewidth=0.95) +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=2) +
  geom_sf(data=res_pts,aes(colour="Beaver Lake #131 Reserve"),fill="blue",linewidth=1.5) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,11,15)),size=2)+
geom_text_repel(
    data = larp_pop %>% filter(OBJECTID %in% c(3, 6, 7, 8, 11, 15)) %>% 
      st_transform(crs = 4326),  # or match the CRS of your plot
    aes(label = NAME, geometry = geometry),
    stat = "sf_coordinates",
    nudge_y = 10,  # degrees if using EPSG:4326, meters otherwise
    size = 5
  )+
  scale_colour_manual("",values = c("blue","firebrick","orange"))+
  coord_sf(
    xlim = c(bbox_coords$xmin, bbox_coords$xmax),
    ylim = c(bbox_coords$ymin, bbox_coords$ymax)
  )+
  theme_minimal()+
  #scale_fill_manual("")+
 #theme_void()+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  labs(x="",y="",
      title="Crude Oil Pipeline Infrastructure Within BLCN Territory",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    



```






```{r proj_map,fig.align='center',fig.width=16, fig.height=8,dpi=220}



osb_1980 <- st_read("../blcn/BLCN OSProject Boundaries1980s.gpkg",quiet = TRUE)%>%clean_names()

osb_1998 <- st_read("../blcn/BLCN OSProject Boundaries1998.gpkg",quiet = TRUE)%>%clean_names()

osb_2007 <- st_read("../blcn/BLCN OSProject Boundaries2007.gpkg",quiet = TRUE)%>%clean_names()

osb_2011 <- st_read("../blcn/BLCN OSProject Boundaries2011.gpkg",quiet = TRUE)%>%clean_names()

osb_2015 <- st_read("../blcn/BLCN OSProject Boundaries2015.gpkg",quiet = TRUE)%>%clean_names()

download.file("https://www.aer.ca/data/shapefiles/Scheme_Approval_SHP.zip",destfile = "ab_regimes.zip",mode="wb")
unzip("ab_regimes.zip")

osb<-bind_rows(osb_1980,osb_1998,osb_2007,osb_2015)%>%clean_names()%>%
  group_by(osp_no)%>%
  mutate(first_year=as.numeric(min(year)),
         project_name=last(project_name))%>%
  ungroup()%>%
  arrange(operator_name,project_name)%>%
  st_transform(3347)

#sf_use_s2(FALSE)


in_situ_shp <- st_read("Shapefiles/In Situ Oil Sands Scheme Approvals/In Situ Oil Sands Scheme Approval_GCS_NAD83.shp",quiet=TRUE)%>%clean_names()%>%
  st_transform(crs=st_crs(gd_pts))
  

#AER ST 102 Facilities List

download.file("https://www.aer.ca/data/codes/ST102-SHP.zip","ST102.zip",mode="wb")
unzip("ST102.zip")

ST102<-st_read("ST102_Facility_GCS_NAD83.shp",quiet=TRUE)%>%clean_names()%>%
  st_transform(crs=st_crs(gd_pts))

#AER ST44 Oil sands Schemes Excel
  
download.file("https://www.aer.ca/prd/documents/sts/ST44_2024-Q4.xlsx",dest="aer_os_schemes.xlsx",mode="wb")

aer_active_schemes<-read_excel("aer_os_schemes.xlsx",sheet="ActiveInSitu",skip=3)%>%clean_names()%>%
  rename(sche_name=scheme_name)%>%
  mutate(active=TRUE)

aer_inactive_schemes<-read_excel("aer_os_schemes.xlsx",sheet="Inactive",skip=3)%>%clean_names()%>%
  rename(sche_name=scheme_name)%>%
  mutate(active=FALSE)

aer_schemes<-bind_rows(aer_active_schemes,aer_inactive_schemes)%>%
  mutate(scheme_num=str_extract(approval_number, "^\\d+"))%>%
  distinct()
              
  
in_situ <- st_filter(in_situ_shp, blcn_terr, .predicate = st_intersects)%>%
    mutate(scheme_num=str_extract(scheme_no, "^\\d+"),
           sche_name = replace(sche_name, is.na(sche_name), ""))%>%
    left_join(aer_schemes%>%
        mutate(sche_name = replace(sche_name, is.na(sche_name), "")),
        by=c("scheme_num","sche_name"))%>%
    mutate(status=ifelse(is.na(status),"",status))#%>%
    #filter(status!="Expired")%>%
    #filter(status!="Terminated Not Started")%>%
    #filter(status!="Superseded")%>%
    #mutate(test_st53=scheme_num%in% ST53_appr$scheme_num)
  
#exp<-in_situ %>% select(scheme_no,scheme_num,scheme_typ,approval_number,app_holder,area)

#save(exp,file="../st53/blcn_app.rdata")





#     mutate(issue_group = case_when(
#      issue_date < as.Date("2000-01-01") ~ "pre-2000",
#      issue_date >= as.Date("2000-01-01") & issue_date <= as.Date("2008-12-31") ~ #"2000–2008",
#      issue_date >= as.Date("2009-01-01") & issue_date <= as.Date("2015-12-31") ~ #"2009–2015",
#      issue_date >= as.Date("2016-01-01") ~ "2016–present",
#    TRUE ~ NA_character_
#  ))
    
  

  in_situ_oth<-in_situ %>%
    filter(sch_status%in%c("AC","TS","AC","TU","TN","SU"))%>%
    filter(sch_sub_ty %in% c("Primary","Commercial","Enhanced Recovery"))

#write.csv(file = "projects_osip.csv",osb%>%st_drop_geometry()%>% #select(project_name,operator_name)%>%distinct())

```

```{r}
ggplot(osb) +
  geom_sf(aes(fill = as_factor(first_year)), color = "white") +
  #geom_sf(data=in_situ%>%mutate(year=2024),aes(fill = "Current"), color = "white") +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =             #                                                                     NAME),nudge_y=.50)+
  geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =                                                                                 NAME),nudge_y=10000)+
  scale_fill_viridis("",begin = 0.8,end = 0.1,discrete=T)+
  scale_color_brewer("",palette = "Set1")+
 tombe_theme()+
  theme(
    legend.position = "bottom",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  facet_wrap(~year,nrow = 1)+
  labs(x="",y="",title="Oil Sands Project Boundaries Within BLCN Territory Over Time",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    
#"~/data_projects/blcn/Great Divide project boundary 2015.gpkg"

```


```{r}


 company_levels <- c(
  "Harvest", "IPC", "Cenovus", "MEG", "Imperial",
  "Athabasca", "CNRL", "Grizzly", "Strathcona",
  "ConocoPhillips",                      # main entries
  "Other Primary and EOR Regimes"       # must come last
)



in_situ_combo<-in_situ%>%
    mutate(sche_name=ifelse(is.na(sche_name),"",sche_name))%>%
    group_by(sche_name,app_holder) %>%
    summarize(geometry = st_union(geometry),
              sch_sub_ty=last(sch_sub_ty),
              .groups = "drop")%>%
    I()%>%
  mutate(sche_name=ifelse(sche_name=="Christina Lake (MEG)","Christina Lake",sche_name))%>%
  mutate(app_holder=case_when(
     app_holder=="Harvest Operations Corp." ~ "Harvest",
     app_holder=="Ipc Canada Ltd." ~"IPC",                      
     app_holder=="Cenovus Energy Inc."~"Cenovus",
     app_holder=="MEG Energy Corp."~"MEG",
     app_holder=="Imperial Oil Resources Limited"~"Imperial",
     app_holder=="Athabasca Oil Corporation"~"Athabasca",
     app_holder=="Canadian Natural Resources Limited"~"CNRL",
     app_holder=="Grizzly Oil Sands ULC"~"Grizzly",
     app_holder=="Strathcona Resources Ltd."~"Strathcona",
     app_holder=="ConocoPhillips Canada Resources Corp."~"ConocoPhillips"
  ))%>%
  mutate(label=paste(sche_name," (",app_holder,")",sep=""))%>%
  mutate(app_holder = factor(app_holder, levels = company_levels))
  
in_situ_oth<-in_situ_oth%>%
  mutate(app_holder=case_when(
     app_holder=="Cenovus Energy Inc."~"Cenovus",
     app_holder=="Canadian Natural Resources Limited"~"CNRL",
     app_holder=="Strathcona Resources Ltd."~"Strathcona",
     TRUE~"Other Primary and EOR Regimes"
  ))%>%
  mutate(app_holder = factor(app_holder, levels = company_levels))

palette<-c(colors_tableau10(),colors_tableau10_light())

ggplot(in_situ_combo)+
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  #geom_sf(data=res_pts,aes(colour="Beaver Lake 131 Reserve"),fill=NA,linewidth=1) +
  geom_sf(data=clawr_data,aes(colour="Cold Lake Air Weapons Range"),fill=NA,linewidth=1) +
  geom_sf(aes(colour="Oil Sands Scheme Approval Boundaries",fill=app_holder),linewidth=0.5) +
  geom_sf(data=in_situ_oth,aes(colour="Oil Sands Scheme Approval Boundaries",fill=app_holder),linewidth=0.5)+
  geom_label_repel(
    aes(label = label,geometry=geometry),size=2.5, color="black",
    stat = "sf_coordinates",
    max.overlaps = Inf,
    min.segment.length = 0
  )+
  
  #geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
  scale_color_manual("",values=c("dodgerblue","darkgreen","black"))+
  scale_fill_manual("",
  values = c(
    "Harvest" = palette[1],
    "IPC" = palette[2],
    "Cenovus" = palette[3],
    "MEG" = palette[4],
    "Imperial" = palette[5],
    "Athabasca" = palette[6],
    "CNRL" = palette[7],
    "Grizzly" = palette[8],
    "Strathcona" = palette[9],
    "ConocoPhillips" = palette[10],
    "Other Primary and EOR Regimes" = palette[11]
  )
)+
  tombe_theme()+
  guides(colour=guide_legend(nrow = 3))+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  #facet_wrap(~year,nrow = 1)+
  labs(x="",y="",title="Oil Sands Projects Within BLCN Territory",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
  
ggsave("test.pdf",width=16,height = 9,dpi=300)

```


```{r} 
## Wells

st37_txt <- read_tsv("../blcn/WellList.txt",col_names = FALSE)%>%
  setNames(c(
  "UWI ID",
  "Loc-Desc",
  "Update-Flag",
  "Well-Name",
  "Field Code",
  "Pool Code",
  "OS-Area- Code",
  "OS-Dep- Code",
  "License-No",
  "License-Status",
  "License-Issue-Date",
  "Licensee-Code",
  "Agent-Code",
  "Operator-Code",
  "Fin-Drl-Date",
  "Well-Total-Depth",
  "Well-Stat-Code",
  "Well-Stat-Date",
  "Fluid_Short_Description",
  "Mode_Short_Description",
  "Type_Short_Description",
  "Structure_Short_Description",
  "Scheme_Type",
  "Scheme_Sub_Type"
))%>%clean_names()





 st37_shp <- st_read("../blcn/ST37_SH/ST37_SH_GCS_NAD83.shp")%>%clean_names()%>%
  st_transform(crs=st_crs(gd_pts))
  
st37_dh <- st_read("../blcn/ST37_BH/ST37_BH_GCS_NAD83.shp")%>%clean_names()%>%
  st_transform(crs=st_crs(gd_pts))


  
st37_blcn<- st_intersection(st37_shp,blcn_buffer)

st37_dh_blcn<- st_intersection(st37_dh,blcn_buffer)

st37_blcn<- st37_blcn %>% group_by(licence,lic_status)%>%mutate(records=n())%>% ungroup()
  


st37_blcn<-st37_blcn%>%left_join(st37_txt%>%select(licence=license_no,fluid_short_description,type_short_description,fin_drl_date)%>%distinct())%>%
  filter(fluid_short_description!="WATER")%>%
  mutate(lic_stat_dat=dmy(lic_stat_dat),
         year=round(year(lic_stat_dat),-1))%>%
  group_by(licence)%>%
  mutate(first=pmin(lic_stat_dat),
         last=pmax(lic_stat_dat))%>%
  ungroup()%>%
  I()
  

non_op<-c("RecExempt","RecCertified","Abandoned","Suspension")

st37_blcn<-st37_blcn%>%
  group_by(licence)%>%
  arrange(lic_stat_dat)%>%
  mutate(last_status=last(lic_status),last_date=last(lic_stat_dat))%>%
  ungroup()%>%
  mutate(op=(!last_status%in%non_op))%>%
  mutate(status=case_when(
    op ~ "Operating at present",
    last_status=="Suspension" ~ "Reclaimed, Abandoned or Suspended since 2008",
    last_status=="Abandoned" ~ "Reclaimed, Abandoned or Suspended since 2008",
    last_status%in%c("RecExempt","RecCertified") ~ "Reclaimed, Abandoned or Suspended since 2008",
    TRUE~"Other"
    ))%>%
  mutate(relevant=(op|last_date>=ymd("2008-01-01")))
    
    
  
    



ggplot(st37_blcn%>%filter(relevant,fluid_short_description %in% c("GAS","CR-BIT","CR-OIL")|type_short_description=="SAGD")) +
  geom_sf(aes(colour=status),size=0.3)+
  geom_sf(data=osb,aes(color = "Oil Sands Project Boundaries"),fill=NA,linewidth=0.5)+
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
  geom_sf(data=res_pts,aes(colour="Beaver Lake Cree Nation Reserve"),fill=NA,linewidth=1) +
    #geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)))+
scale_color_manual("",values=colors_tableau10())
  


st37_dh_bl<- st37_dh_blcn %>% mutate(fd_date_n=ymd(fd_date))

```


```{r}
aepa_projects <- read_excel("../blcn/epa-oil-sands-operating-capacity-report-2023-2024.xlsx",sheet="2022 (BY PHASE)") %>% clean_names()

test<-aepa_projects %>% 
  #mutate(project_name=ifelse(grepl("Cold Lake"),"Cold",project_name))%>%
  select(operator,project_name) %>% distinct()
  
  
```



## ST-53


```{r}

load("../st53/st53_all.rdata")

load("blcn_app.rdata")

blcn_projects<-c(
        #"Blackrod",
        "BlackGold",
        "Christina Lake (CVE)",
        "Christina Lake (MEG)",
        "Cold Lake",
        "Foster Creek",
        "Jackfish",
        "Kirby",
        "Leismer",
        "Primrose and Wolf Lake",
        "Burnt Lake",
        #"Narrows Lake",
        "Orion",
        "Pike 1",
        "Surmont",
        "Tucker",
        #"May River",
        "CNRL Primary"
        )

options(scipen = 999)
BLCN<-ST53_data%>%clean_names()%>%
  mutate(scheme_num=str_extract(approval_number, "^\\d+"))%>%
  filter(scheme_num %in% exp$scheme_num)%>%
  mutate(project=
         if_else(is.na(scheme_name),paste(operator,recovery_method),scheme_name),
         project=gsub("Canadian Natural Resources Limited","CNRL",project),year=year(date)
         )%>%
  mutate(monthly_prod=bitumen*days_in_month(date))%>%
  group_by(project,year)%>%
  mutate(annual_prod=sum(monthly_prod,na.rm=T)*6.2898/10^6)%>%
  ungroup()%>%
  arrange(-annual_prod)%>%
  mutate(project=as_factor(project),
        )

top_9<- levels(BLCN$project)[1:9]

BLCN<-BLCN%>%mutate(
  project=fct_other(project,keep=top_9,other_level="Other Projects")
)  %>%
  group_by(project,date)%>%
  summarize(bitumen=sum(monthly_prod,na.rm = T))%>%
  ungroup()
  
  
BLCN%>%
      ggplot(aes(date,bitumen))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  scale_x_continuous(expand=c(0,0),breaks=pretty_breaks(6))+
  scale_y_continuous(expand=c(0,0),breaks=pretty_breaks(5))+
  #coord_flip()+
  #scale_fill_brewer()+
  #guides(fill="none",color="none")+
  facet_wrap(~project,nrow=2)+
  #scale_x_reverse()+
  theme_ps_grid()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Bitumen Production (million barrels per year)",
       title="Annual Bitumen Production by Project, Beaver Lake Cree Nation Core Territory",
       #subtitle=paste("Total production since 2016 of ",round(total$prod,2)," billion barrels and 2023 average production of ",sum(blcn_data%>%filter(reporting_year==2023)%>%select(prod)/365/10^6)%>%round(2)," million barrels per day",sep=""),
       caption="Data via Government of Alberta, graph by @andrew_leach")  
  
```

```{r}

#Sproule Valuation
library(rvest)
# Load the Sproule price forecast page
url <- "https://sproule.com/price-forecast/"
page <- read_html(url)

# Extract all links
all_links <- page %>%
  html_elements("a") %>%
  html_attr("href") %>%
  na.omit()

# Filter for links that likely point to the Escalated Forecast Excel
forecast_links <- all_links %>%
  str_subset("(?i)escalated.*\\.xlsx$")  # case-insensitive match for .xlsx files containing "escalated"

download.file(first(forecast_links),destfile = "sproule_current.xlsx",mode="wb")

sproule_hist<-read_excel("sproule_current.xlsx",sheet=" History",skip=9)%>%
  clean_names()%>%
  mutate(month=excel_numeric_to_date(as.numeric(month)))%>%
  filter(!is.na(month))%>%
  rename(date=month)%>%
  mutate(across(-date, ~ as.numeric(.)))%>%
  filter(!if_all(-date, is.na))

header_rows <- read_excel("sproule_current.xlsx",sheet=" History",skip=6,col_names = FALSE)%>%
  slice(1:3)


header_rows <- read_excel("sproule_current.xlsx",
                          sheet = " History",
                          skip = 6,
                          n_max = 3,      # Only get 3 rows
                          col_names = FALSE)




# Check it's actually a data frame
if (!is.data.frame(header_rows)) stop("Header rows not read correctly.")
#create names

collapsed_names <- map_chr(seq_len(ncol(header_rows)), function(i) {
  parts <- as.character(header_rows[[i]])
  parts <- parts[!is.na(parts)]
  if (length(parts) == 0) {
    return(paste0("V", i))  # fallback name if all NA
  } else {
    return(paste(parts, collapse = "_"))
  }
}) %>%
  str_replace_all("_+", "_") %>%
  str_remove("^_") %>%
  str_remove("\\*+") %>%
  str_remove("_$")%>%
  str_replace_all("_"," ")
  


collapsed_names<-c("date",collapsed_names)
# Apply the new names
names(sproule_hist) <- collapsed_names


unit_row <- read_excel("sproule_current.xlsx",
                          sheet = " History",
                          skip = 9,
                          n_max = 1,      # Only get 1 rows
                          col_names = FALSE)%>%
  as.character()

# Create the 2-column tibble
name_unit_df <- tibble(
  commodity = collapsed_names,
  unit = unit_row
)



sproule_hist<-sproule_hist %>% pivot_longer(-date,names_to = "commodity",values_to = "price")%>%left_join(name_unit_df,relationship = "many-to-many")%>%
  filter(!is.na(price))


BLCN<-BLCN %>%
  mutate(date=floor_date(date,"month"))%>%
  left_join(sproule_hist%>%filter(commodity=="Hardisty Heavy 12 API"))%>%
  mutate(value=bitumen*price)

```

