---
title: "Alberta Oil Sands Royalty Data"
author: Andrew Leach
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      code_folding: hide
      includes:
      after_body: 
      theme: lumen
always_allow_html: yes
editor: source

---
```{r, echo=FALSE,include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,cache = T,fig.align='center',dpi=220,fig.width=16,fig.height=9)
options(scipen = 999)
```

One of the recommendations of the Alberta Royalty Review in 2015-16 was more transparency with respect to oil sands costs, profits, production and royalty payments. The Alberta Government has followed that recommendation and makes [project-level data](https://open.alberta.ca/opendata/alberta-oil-sands-royalty-data1) available on their open data site. This document compiles those data and provides some basic graphs that you might find useful. Where appropriate, I've provided a link to the R code used to process the data.

```{r,cache=F}
library(janitor)
library(scales)
library(tidyverse)
library(readxl)
library(openxlsx)
library(viridis)
library(ggthemes)
library(kableExtra)

work_theme<-function(){
  theme_tufte(18)+
     theme(
           plot.subtitle = element_text(color="grey10",size=rel(1)),
           plot.title = element_text(face="bold"),
           plot.caption = element_text(color="grey50",size=rel(1)),
           legend.title = element_text(color="grey10",size=rel(1.5)),
           legend.text = element_text(color="grey10",size=rel(1.5)),
           strip.text = element_text(size=rel(1.2)),
           axis.title = element_text(size=rel(1.2)),
           axis.text = element_text(size=rel(1.2)),
           axis.text.x = element_text(size=rel(.8)),
           axis.ticks = element_blank(),
           panel.spacing = unit(.75,"lines"),
           legend.position = "bottom",
           plot.margin = margin(t = .5, r = .5, b = .5, l = .5,unit= "cm"),
           #axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)
           NULL
           )+
  NULL
}

colors_tableau10<-function () 
{
    return(c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", 
        "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17BECF"))
}

```

## Download and Combine the Annual Data for 2016-2021

The first step is to load and process the data. The code button provides a window into how I've done this.


```{r load data,cache=T}
#load plant data
make_os_data<-function(){
os_data_2021 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2021 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2020 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2020 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  
  os_data_2019 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2019 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
  os_data_2018 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2018 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

os_data_2017 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2017 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
os_data_2016 <- read.xlsx(xlsxFile = "royalty_data.xlsx", sheet = "2016 Royalty Data", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)

#stack all the years together

os_data<-bind_rows(os_data_2021,os_data_2020,os_data_2019,os_data_2018,os_data_2017,os_data_2016)

#clean the names

os_data<-os_data %>% clean_names()%>%
  mutate(op_profit=gross_revenue-operating_costs-other_costs+other_net_proceeds)


# data are reported for an individual project twice if it passes payout in a given year

# see here for a test
#os_data %>% group_by(project,reporting_year) %>% select(project,project_name,reporting_year,payout_status) %>% mutate(n_obs=n())%>% filter(n_obs>1) %>% arrange(-n_obs)

# combine the data and include two things: a payout year indicator and a net and gross rev royalty number along with a total royalty paid
os_data <- os_data %>% group_by(project,reporting_year,project_name,operator_name)%>%
  summarize(
    project_revenue=sum(project_revenue),
    gross_revenue=sum(gross_revenue),
    op_profit=sum(op_profit),
    cleaned_crude_bitumen_at_rcp_barrels=sum(cleaned_crude_bitumen_at_rcp_barrels),
    operating_costs=sum(operating_costs),
    capital_costs=sum(capital_costs),
    return_allowance=sum(return_allowance),
    other_costs=sum(other_costs),
    other_net_proceeds=sum(other_net_proceeds),
    net_revenue=sum(net_revenue),
    royalty_type=last(royalty_type),
    royalty_type_start=first(royalty_type),
    payout_start=first(payout_status),
    payout_status=last(payout_status),
    net_rev_royalty=sum(royalty_payable*(royalty_type=="NET")),
    gross_rev_royalty=sum(royalty_payable*(royalty_type=="GROSS")),
    royalty_payable=sum(royalty_payable),
    royalty_rate=last(royalty_rate_percent),
    first_royalty_rate=first(royalty_rate_percent),
    royalty_avg=royalty_payable/gross_revenue,
    payout_year=(n()>1),
    unrecovered_balance_net_loss_at_eop=last(unrecovered_balance_net_loss_at_eop))%>%
    mutate(
      last_prod=last(cleaned_crude_bitumen_at_rcp_barrels)/365,
      label=paste(project_name,"\n(",formatC(round(last_prod,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""))%>%
  ungroup()%>%
  group_by(project)%>%
  mutate(project_payout=(last(payout_status)=="POST"))

#fix facilities
os_data <- os_data %>% mutate(
  royalty_bbl=royalty_payable/cleaned_crude_bitumen_at_rcp_barrels,
  op_costs_bbl=operating_costs/cleaned_crude_bitumen_at_rcp_barrels,
  gross_revenue_bbl=gross_revenue/cleaned_crude_bitumen_at_rcp_barrels,
  cap_costs_bbl=capital_costs/cleaned_crude_bitumen_at_rcp_barrels,
  project_name=gsub(" Project","",project_name),
  project_name=gsub("Christina Lake Regional","Christina Lake (MEG)",project_name),
  project_name=gsub("Christina Lake Thermal","Christina Lake (CVE)",project_name),
  project_name=gsub("MacKay River Commercial","PetroChina",project_name),
  project_name=gsub("MacKay River","MacKay River (Suncor)",project_name),
  project_name=gsub("PetroChina","MacKay River (PetroChina)",project_name),
  project_name=gsub(" Thermal","",project_name),
  project_name=gsub(" Mine","",project_name),
  project_name=gsub(" Oil Sands","",project_name),
  project_name=gsub(" EOR","",project_name),
  project_name=gsub(" Commercial","",project_name),
  project_name=gsub(" SAGD","",project_name),
  project_name=gsub(" Demonstration","",project_name),
  project_name=gsub(" In-Situ","",project_name),
  project_name=as.factor(project_name),
  op_profit_net=op_profit-royalty_payable,
  op_profit_bbl=op_profit/cleaned_crude_bitumen_at_rcp_barrels,
  op_profit_net_bbl=op_profit_net/cleaned_crude_bitumen_at_rcp_barrels
  )

}

os_data<-make_os_data()

    
mines<-c("Muskeg River","Fort Hills","Kearl","Horizon","Muskeg River","Jackpine","Syncrude","Suncor")

SAGD<-c("Hangingstone","Leismer","Blackrod","Mackay River","Kirby","Christina Lake","Foster Creek",
        "Long Lake","Great Divide","Surmont","Jackfish","Sunrise","Orion","Firebag","MacKay River",
        "West Ells","SAGD","Tucker","Narrows Lake")

os_data<-os_data %>% mutate(mine=(project_name %in% mines),
                            in_situ=(!project_name %in% mines),
                            )%>%
  group_by(project_name)%>%
  mutate(min_bbls=min(cleaned_crude_bitumen_at_rcp_barrels)/365,
         max_bbls=max(cleaned_crude_bitumen_at_rcp_barrels)/365)%>%
  ungroup()%>%
  mutate(project=(max_bbls>8000),
         big_project=(min_bbls>30000),
         med_project=(min_bbls>8000)&(max_bbls<50000)
         )


```

## Production

```{r mines prod,fig.align='center',fig.width=16, fig.height=8,dpi=220}

ggplot(os_data %>% filter(project,mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod)
                ),
       aes(factor(reporting_year),last_prod/1000,fill=project_type))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=-90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Bitumen Production (thousands of barrels per day)",
       title="Annual Bitumen Production, Oil Sands Mining Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

```{r in situ prod,fig.align='center',fig.width=16, fig.height=8,dpi=220}

ggplot(os_data %>% filter(big_project,!mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod)
                ),
       aes(factor(reporting_year),last_prod/1000,fill=project_type))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Bitumen Production (thousands of barrels per day)",
       title="Annual Bitumen Production, Larger Oil Sands In Situ Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```


## Revenue per Barrel


```{r mines rev,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod)
                ),
       aes(factor(reporting_year),gross_revenue_bbl,fill=project_type))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Revenue per barrel bitumen (CA$/bbl)",
       title="Gross Revenue per Barrel Bitumen, Oil Sands Mining Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

```{r in situ rev,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(big_project,!mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod)
                ),
       aes(factor(reporting_year),gross_revenue_bbl,fill=project_type))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Revenue per barrel bitumen (CA$/bbl)",
       title="Gross Revenue per Barrel Bitumen, Larger In Situ Oil Sands Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

## Operating Costs


```{r mining op costs,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod)
                ),
       aes(factor(reporting_year),op_costs_bbl,fill=project_type))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Operating cost per barrel bitumen (CA$/bbl)",
       title="Operating Costs, Oil Sands Mining Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```


```{r in situ op costs,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(big_project,!mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod)
                ),
       aes(factor(reporting_year),op_costs_bbl,fill=project_type))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  scale_fill_brewer()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Operating cost per barrel bitumen (CA$/bbl)",
       title="Operating Costs, Larger In Situ Oil Sands Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

```{r}
#test<-
  os_data %>% 
  #filter(!mine)%>%
  group_by(project_name)%>%
  mutate(
      last_prod=last(cleaned_crude_bitumen_at_rcp_barrels)/365,
      label=paste(project_name,"\n(",formatC(round(last_prod,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""))%>%
  ungroup()%>%
  filter(last_prod>10000)%>%
  mutate(label=factor(label),
      project_type=ifelse(mine,"Mine","In Situ"),
      label=fct_reorder(label,last_prod),
      op_costs_bbl=format(op_costs_bbl,nsmall=2),
      op_costs_bbl=ifelse(grepl("NaN",op_costs_bbl),"",op_costs_bbl))%>%
    select(label,project_type,reporting_year,op_costs_bbl)%>%
  pivot_wider(names_from = reporting_year,values_from = op_costs_bbl)%>%
  rename("Project"=label,"Project Type"=project_type)%>%
  arrange(Project)%>%
  #rename("2021 Production"=last_prod)%>%
  rename_all( ~ gsub("x","",.))%>%
  kable(table.attr = "style='width:80%;'",align = c("l","c", rep("r", 6)), linesep = "", escape = FALSE) %>% 
  kable_styling(fixed_thead = T,bootstrap_options = c("hover", "condensed","responsive"),full_width = T)%>%
  I() 
  #row_spec(1, bold = T)

```


## Operating Cost Density Plot


```{r,fig.align='center',fig.width=16, fig.height=12,dpi=220}
ggplot(os_data%>%filter(big_project) %>% group_by(reporting_year) %>% 
  mutate(year_total=sum(cleaned_crude_bitumen_at_rcp_barrels), weight=cleaned_crude_bitumen_at_rcp_barrels/year_total) %>%
  ungroup()%>%
    mutate(mine=as.factor(mine),
           mine=fct_recode(mine,Mine="TRUE","In Situ"="FALSE"))
           
           , aes(op_costs_bbl,group=factor(reporting_year),weights=weight))+
  stat_density(aes(color=factor(reporting_year)),geom="line",position = "identity",trim=T,size=1.6)+
  scale_colour_manual(NULL,values=colors_tableau10())+
  facet_wrap(~mine,scales="free_x")+
  expand_limits(x=0)+
  guides(color=guide_legend(nrow=1))+
  #geom_density(aes(color=factor(reporting_year)), alpha=0.8) + 
  labs(title="Density plot of oil sands operating costs per barrel bitumen", 
       subtitle="Production-weighted, for projects with more than 10,000 barrels per day of bitumen production",
       caption="Source: Government of Alberta 2016 and 2017 Royalty Data, graph by Andrew Leach",
       x="Operating Costs ($Cdn/bbl bitumen)",
       fill="Reporting Year")+
  work_theme()
```



## Royalties



```{r mine royalties,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod),
                payout=as_factor(payout_status),
                payout=fct_relevel(payout,"PRE")
                ),
       aes(factor(reporting_year),royalty_bbl,fill=payout))+
  scale_fill_manual("",values=c("PRE"="lightgreen","POST"="darkgreen"),labels=c("Pre-Payout","Post-Payout"))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  guides(fill=guide_legend(nrow = 1),color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Royalties payable per barrel bitumen (CA$/bbl)",
       title="Royalties Payable per Barrel Bitumen, Oil Sands Mining Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```


```{r in situ royalties,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(big_project,!mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod),
                payout=as_factor(payout_status),
                payout=fct_relevel(payout,"PRE")
                ),
       aes(factor(reporting_year),royalty_bbl,fill=payout))+
  scale_fill_manual("",values=c("PRE"="lightgreen","POST"="darkgreen"),labels=c("Pre-Payout","Post-Payout"))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  guides(fill=guide_legend(nrow = 1),color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Royalties payable per barrel bitumen (CA$/bbl)",
       title="Royalties Payable per Barrel Bitumen, Larger In Situ Oil Sands Projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")
```


## Operating Profits (Post-Royalty)

```{r mine net profits,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(mine,project_name!="Fort Hills")%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod),
                payout=as_factor(payout_status),
                payout=fct_relevel(payout,"PRE")
                ),
       aes(factor(reporting_year),op_profit_net_bbl,fill=payout))+
  scale_fill_manual("",values=c("PRE"="lightgreen","POST"="darkgreen"),labels=c("Pre-Payout","Post-Payout"))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  guides(fill=guide_legend(nrow = 1),color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Operating profit per barrel bitumen (CA$/bbl)",
       title="Operating Profit per Barrel Bitumen, Oil Sands Mining Projects",
       subtitle="Gross revenue net operating and capital costs and royalties, excluding Fort Hills",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```


```{r in situ net profits,fig.align='center',fig.width=15, fig.height=8,dpi=220}

ggplot(os_data %>% filter(big_project,!mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod),
                payout=as_factor(payout_status),
                payout=fct_relevel(payout,"PRE")
                ),
       aes(factor(reporting_year),op_profit_net_bbl,fill=payout))+
  scale_fill_manual("",values=c("PRE"="lightgreen","POST"="darkgreen"),labels=c("Pre-Payout","Post-Payout"))+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  #scale_fill_viridis(discrete=FALSE,"Royalties Paid ($/bbl)")+
  #coord_flip()+
  guides(fill=guide_legend(nrow = 1),color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 7, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Operating profit per barrel bitumen (CA$/bbl)",
       title="Operating Profit per Barrel Bitumen, Oil Sands In-Situ Projects",
       subtitle="Gross revenue net operating and capital costs and royalties",
       caption="Data via Government of Alberta, graph by @andrew_leach")
```


```{r}
#test<-
  os_data %>% #filter(!mine)%>%
  
  group_by(project_name)%>%
  mutate(
      last_prod=last(cleaned_crude_bitumen_at_rcp_barrels)/365,
      label=paste(project_name,"\n(",formatC(round(last_prod,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""))%>%
  ungroup()%>%
  filter(last_prod>10000)%>%
  mutate(label=factor(label),
       project_type=ifelse(mine,"Mine","In Situ"),
      label=fct_reorder(label,last_prod),
      op_profit_net_bbl=op_profit_net_bbl/gross_revenue_bbl*100,
      op_profit_net_bbl=format(round(op_profit_net_bbl,2),nsmall=2),
      op_profit_net_bbl=ifelse(grepl("NaN",op_profit_net_bbl),"",op_profit_net_bbl),
      op_profit_net_bbl=ifelse(grepl("Inf",op_profit_net_bbl),"",op_profit_net_bbl)
      )%>%
    select(label,project_type,reporting_year,op_profit_net_bbl)%>%
  pivot_wider(names_from = reporting_year,values_from = op_profit_net_bbl)%>%
  rename("Project"=label,
         "Project Type"=project_type)%>%
  arrange(Project)%>%
  #rename("2021 Production"=last_prod)%>%
  rename_all( ~ gsub("x","",.))%>%
  kable(table.attr = "style='width:80%;'",align = c("l","c", rep("r", 6)), linesep = "", escape = FALSE) %>% 
  kable_styling(fixed_thead = T,bootstrap_options = c("hover", "condensed","responsive"),full_width = T)%>%
  I() 
  #row_spec(1, bold = T)

```
## Unrecovered capital costs

```{r mining cap_cost,fig.align='center',fig.width=14, fig.height=6,dpi=220}

ggplot(os_data %>% filter(project,!project_payout,mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod),
                payout=as_factor(payout_status),
                payout=fct_relevel(payout,"PRE")
                ),
       aes(factor(reporting_year),unrecovered_balance_net_loss_at_eop/10^6,fill=mine))+
  #scale_fill_manual("",values=c("FALSE"="lightgreen","TRUE"="darkgreen"),labels=c("In Situ","Mine"))+
  #scale_fill_brewer("Project type")+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  scale_fill_brewer()+
  #coord_flip()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 1,labeller = label_wrap_gen(width = 10, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Unrecovered capital costs (CA$ million)",
       title="Unrecovered capital costs per royalty formula, oil sands mining projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

```{r in situ cap_cost,fig.align='center',fig.width=16, fig.height=12,dpi=220}
ggplot(os_data %>% filter(project,!project_payout,!mine)%>%
         mutate(project_type=ifelse(mine,"Mine","In Situ"),
                project_name_factor=as.factor(project_name),
                project_name_factor=fct_reorder(project_name_factor,last_prod),
                payout=as_factor(payout_status),
                payout=fct_relevel(payout,"PRE")
                ),
       aes(factor(reporting_year),unrecovered_balance_net_loss_at_eop/10^6,fill=mine))+
  #scale_fill_manual("",values=c("FALSE"="lightgreen","TRUE"="darkgreen"),labels=c("In Situ","Mine"))+
  #scale_fill_brewer("Project type")+
  geom_col(size=0.25,position = position_dodge(width = .25),color="black")+
  #scale_color_viridis("Royalties Paid\nPer Barrel\nCleaned Bitumen")+
  scale_fill_brewer()+
  #coord_flip()+
  guides(fill="none",color="none")+
  facet_wrap(~project_name_factor,nrow = 2,labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
  #scale_x_reverse()+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(x=NULL,y="Unrecovered capital costs (CA$ million)",
       title="Unrecovered capital costs per royalty formula, larger in situ oil sands projects",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, graph by @andrew_leach")

```

## GHG emissions intensity

```{r ei 1,fig.align='center',fig.width=16, fig.height=7,dpi=220}
#load plant data
os_ghg_data <- read_excel(path = "ab_os_ghgs.xlsx", sheet = "Emission Intensity", range="A4:AO35")%>%
  clean_names()%>% pivot_longer(cols=seq(15,23),names_to="year",values_to="adj_ghg")%>%
  select(company,facility,subsector,product,year,adj_ghg)%>%mutate(year=str_sub(year,start=2,end=5))

os_prod_data <- read_excel(path = "ab_os_ghgs.xlsx", sheet = "Emission Intensity", range="A4:AO35")%>%
  clean_names()%>% pivot_longer(cols=seq(24,32),names_to="year",values_to="prod")%>%
  select(company,facility,subsector,product,year,prod)%>%mutate(year=str_sub(year,start=2,end=5))

os_ei_data <- read_excel(path = "ab_os_ghgs.xlsx", sheet = "Emission Intensity", range="A4:AO35")%>%
  clean_names()%>% pivot_longer(cols=seq(33,41),names_to="year",values_to="ei")%>%
  select(company,facility,subsector,product,year,ei)%>%mutate(year=str_sub(year,start=2,end=5)) %>% 
  left_join(os_ghg_data)%>% 
  left_join(os_prod_data)



#keep anything larger than tucker
os_big_projects<- os_ei_data %>% filter(year==2019) %>% group_by(subsector)%>% filter(prod>1250000)

os_ei_data %>% mutate(facility=factor(facility),facility=fct_other(facility,keep = os_big_projects$facility))%>%
  group_by(facility,subsector,product,year) %>% summarize(prod=sum(prod),ei=sum(adj_ghg)/sum(prod),adj_ghg=sum(adj_ghg))%>%
  mutate(facility=fct_recode(facility, "Foster Creek" = "Foster Creek SAGD Bitumen Battery (with Cogen)",
                             "Canadian Natural AOSP"="Canadian Natural Upgrading Limited Muskeg River Mine and Jackpine Mine and Scotford Upgrader",
                             "MEG Christina Lake"="MEG Christina Lake Regional project",
                             "Cenovus Christina Lake"="Christina Lake SAGD Bitumen Battery",
                             "Hangingstone"="Hangingstone Expansion project",
                             "MacKay River"="MacKay River, In-Situ Oil Sands Plant",
                             "Surmont"="Surmont SAGD Commercial Battery"))%>%
  group_by(facility) %>% mutate(prod_2019=sum(prod*(year==2019)*6.2929/365))%>%
  mutate(label2=paste(facility,"\n(",formatC(round(prod_2019,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""),
         label=facility)%>%
  filter(subsector!="In Situ",ei<2)%>%
  ggplot()+
  geom_col(aes(year,ei,fill="A"),size=0.25,position = position_dodge(width = .25),color="black")+
  scale_fill_brewer()+
  #geom_hline(aes(yintercept = 32.65*1.26,colour="Current $CA bitumen value"))+
  facet_wrap(~label,nrow = 1,labeller = label_wrap_gen(width = 12, multi_line = TRUE))+
  #scale_x_reverse()+
  #coord_flip()+
  #scale_fill_viridis("Reporting Year",discrete = T)+
  guides(colour=guide_legend(),fill=FALSE)+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(y="Cogen-adjusted emissions per barrel (t/bbl)",x=NULL,
       title="2011-2019 Emissions Intensity by Oil Sands Mining project",
       #subtitle="projects with production above 40,000 bbl/d in every year. 2019 production, rounded to the nearest thousand barrels per day, shown in brackets.",
       caption="Source: Alberta Government data, graph by @andrew_leach")
```

```{r ei 2,fig.align='center',fig.width=14, fig.height=16,dpi=220}



os_ei_data %>% mutate(facility=factor(facility),facility=fct_other(facility,keep = os_big_projects$facility))%>%
  group_by(facility,subsector,product,year) %>% summarize(prod=sum(prod),ei=sum(adj_ghg)/sum(prod),adj_ghg=sum(adj_ghg))%>%
  mutate(facility=fct_recode(facility, "Foster Creek" = "Foster Creek SAGD Bitumen Battery (with Cogen)",
                                                "Canadian Natural AOSP"="Canadian Natural Upgrading Limited Muskeg River Mine and Jackpine Mine and Scotford Upgrader",
                                                "MEG Christina Lake"="MEG Christina Lake Regional project",
                                                "Cenovus Christina Lake"="Christina Lake SAGD Bitumen Battery",
                                                "Hangingstone"="Hangingstone Expansion project",
                                                "MacKay River"="MacKay River, In-Situ Oil Sands Plant",
                                                "Surmont"="Surmont SAGD Commercial Battery"))%>%
  group_by(facility) %>% mutate(prod_2019=sum(prod*(year==2019)*6.2929/365))%>%
    mutate(label2=paste(facility,"\n(",formatC(round(prod_2019,digits=-3),format="f", big.mark=",",digits = 0)," bbl/d)",sep=""),
         label=facility)%>%
  filter(subsector=="In Situ",ei<2)%>%
  mutate(facility=fct_relevel(facility,after = Inf))%>%
  ggplot()+
  geom_col(aes(year,ei,fill="A"),size=0.25,position = position_dodge(width = .25),color="black")+
  scale_fill_brewer()+
  #geom_hline(aes(yintercept = 32.65*1.26,colour="Current $CA bitumen value"))+
  facet_wrap(~label,nrow = 3,labeller = label_wrap_gen(width = 12, multi_line = TRUE))+
  scale_colour_manual("",values="black")+
  #scale_x_reverse()+
  #coord_flip()+
  #scale_fill_viridis("Reporting Year",discrete = T)+
  guides(colour=guide_legend(),fill=FALSE)+
  work_theme()+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle=90,hjust = 0.5,vjust=0.5)
        )+
  labs(y="Cogen-adjusted emissions per barrel (t/bbl)",x=NULL,
       title="2011-2019 Emissions Intensity by Oil Sands In-Situ project",
       #subtitle="projects with production above 40,000 bbl/d in every year. 2019 production, rounded to the nearest thousand barrels per day, shown in brackets.",
       caption="Source: Alberta Government data, graph by @andrew_leach")

```
