---
title: "Alberta Oil Sands Royalty Data"
author: Andrew Leach
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
      code_folding: hide
      includes:
      after_body: 
      theme: lumen
always_allow_html: yes
editor: source

---
```{r, echo=FALSE,include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,cache = T,fig.align='center',dpi=220,fig.width=16,fig.height=9)
options(scipen = 999)
```

One of the recommendations of the Alberta Royalty Review in 2015-16 was more transparency with respect to oil sands costs, profits, production and royalty payments. The Alberta Government has followed that recommendation and makes [project-level data](https://open.alberta.ca/opendata/alberta-oil-sands-royalty-data1) available on their open data site. This document compiles those data and provides some basic graphs that you might find useful. Where appropriate, I've provided a link to the R code used to process the data.

```{r,cache=F, include=F}
library(janitor)
library(scales)
library(tidyverse)
library(readxl)
library(openxlsx)
library(viridis)
library(ggthemes)
library(kableExtra)
library(sf)

work_theme<-function(){
  theme_tufte(18)+
     theme(
           plot.subtitle = element_text(color="grey10",size=rel(1.25)),
           plot.title = element_text(face="bold"),
           plot.caption = element_text(color="grey50",size=rel(1)),
           legend.title = element_text(color="grey10",size=rel(1)),
           legend.text = element_text(color="grey10",size=rel(1)),
           strip.text = element_text(size=rel(1.2)),
           axis.title = element_text(size=rel(1.2)),
           axis.text = element_text(size=rel(1.2)),
           axis.text.x = element_text(size=rel(.8)),
           axis.ticks = element_blank(),
           panel.spacing = unit(.75,"lines"),
           legend.position = "bottom",
           plot.margin = margin(t = .5, r = .5, b = .5, l = .5,unit= "cm"),
           #axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)
           NULL
           )+
  NULL
}

colors_tableau10<-function () 
{
    return(c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", 
        "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17BECF"))
}

```

<!-- ## Download and Combine the Annual Data for 2016-2022 -->

<!-- The first step is to load and process the data. The code button provides a window into how I've done this. In this case, I've manually combined the data in my own spreadsheet so I can see any changes and also look up individual project information easily. -->


```{r load data,cache=T,include=F}
    file_name<-"https://open.alberta.ca/dataset/3021d788-d4c9-4f49-a849-47b6c02f49cd/resource/ceebbfe4-51b7-4bb6-b4ff-cc694ca09d5b/download/alberta-oil-sands-greenhouse-gas-emission-intensity-analysis.xlsx"
    local_file<-"os_ghgs.xlsx"
    skip_file<-TRUE
      if (file.exists(local_file)) {
      # Get file information
      file_info <- file.info(local_file)
      # Get the last modification time
      mod_time <- file_info$mtime
      # Check if the file was modified in the last 30 days
      if (difftime(Sys.time(), mod_time, units = "days") <= 30) {
      skip_file<-FALSE
      message("The file exists and was created/modified in the last day so not downloading.")
      } else {
        message("The file exists but was not created/modified in the last day.")
      }
    } else {
      message("The file does not exist.")
    }
    if(skip_file)
    download.file(file_name,local_file,mode="wb")
    prod_data<-read_excel(local_file,sheet = "In Situ Production")
    emissions_data<-read_excel(local_file,sheet = "Emissions and Cogen Data")
    ei_data<-read_excel(local_file,sheet = "Emission Intensity",range="A4:BA36")



download.file("",destfile = ,mode = "wb")

```


```{r}
library(httr)
url <- "https://geospatial.alberta.ca/titan/rest/services/boundary/military_base_and_training_ground/MapServer/0/query"

# Set up the query parameters
params <- list(
  where = "1=1",
  outFields = "*",
  f = "geojson"
)

# Make the GET request
response <- GET(url, query = params)

# Check if the request was successful
if (status_code(response) == 200) {
  # Parse the GeoJSON content
  geojson_content <- content(response, as = "text", encoding = "UTF-8")
  clawr_data <- st_read(geojson_content)
} else {
  stop("Failed to retrieve data from the REST service.")
}


```

```{r}
aepa_projects <- read_excel("../blcn/epa-oil-sands-operating-capacity-report-2023-2024.xlsx",sheet="2022 (BY PHASE)") %>% clean_names()

test<-aepa_projects %>% 
  #mutate(project_name=ifelse(grepl("Cold Lake"),"Cold",project_name))%>%
  select(operator,project_name) %>% distinct()
  
  
```




```{r,fig.align='center',fig.width=16, fig.height=8,dpi=220}
all_projects <- st_read("../blcn/OSProjects2015.gdb")%>%clean_names()%>%
  st_transform(crs=st_crs(gd_pts))%>% st_make_valid()%>%
  filter(!project_name%in%c("Caribou","Grouse","Walleye","Birchwood",
                            "Chard","Winefred Lake","West Kirby",
                            "Mariana - Thornbury","South Leismer","Sage"))%>% #cancelled
  filter(!project_name%in%c("Whitesands","Foster Creek North","Jackfish East"))%>%
  mutate(project_name=gsub("Kirby Expansion Project","Kirby",project_name))%>%
  mutate(project_name=gsub("Tucker Lake","Tucker",project_name))%>%
  mutate(project_name=gsub("Cold Lake Operations","Cold Lake (Imperial Oil)",project_name))%>%
  filter(!(operator_name=="MEG Energy Corp."&project_name=="Surmont"))%>%
  filter(!(operator_name=="MEG Energy Corp."&project_name=="May River")) #cancelled


blcn_buffer_20<-st_buffer(blcn_terr, units::as_units(20, 'kilometer'))%>%st_make_valid()%>%
  st_transform(crs=st_crs(gd_pts))

os_blcn_strict<-st_intersection(all_projects,blcn_terr)

os_blcn<-st_intersection(all_projects,blcn_buffer_20)%>%
  filter(project_name %in% os_blcn_strict$project_name)%>%
  #mutate(label=paste(project_name,"\n(",operator_name,")",sep=""))
  mutate(label=project_name)



library(ggrepel)
#test<-"tar"

ggplot(blcn_buffer_20) +
  geom_sf(data=blcn_terr,aes(colour="Beaver Lake Cree Nation Core Territory"),fill=NA,linewidth=1) +
#  geom_sf(data=all_projects %>% filter(grepl(test, project_name)),aes(fill = as_factor(project_name)), color = "white") +
 # geom_sf_label(data = all_projects %>% filter(grepl(test, project_name)), aes(label =project_name),size=1)+
geom_sf(data=os_blcn,aes(),fill="grey40", color = "white") +
  ggrepel::geom_label_repel(
    data = os_blcn,
    aes(label = label, geometry = Shape),size=3.5,
    stat = "sf_coordinates",
    min.segment.length = 0
  )+
  ggrepel::geom_label_repel(
    data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)),
    aes(label = NAME,geometry=geometry),size=3.5, color="red",
    stat = "sf_coordinates",
    min.segment.length = 10,
    nudge_y=-.05
  )+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =                        #                                                         NAME),nudge_y=.10,color="blue")+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =             #                                                                     NAME),nudge_y=.50)+
  #geom_sf_label(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7)), aes(label =                        #                                                         NAME),nudge_y=10000)+
  scale_fill_viridis("",begin = 0.8,end = 0.1,discrete=T,guide=FALSE)+
  scale_color_brewer("",palette = "Set1")+
 tombe_theme()+
  theme(
    legend.position = "bottom",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  #facet_wrap(~year,nrow = 1)+
  labs(x="",y="",title="Oil Sands Projects Within BLCN Territory",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    
#"~/data_projects/blcn/Great Divide project boundary 2015.gpkg"

```



```{r proj_map,fig.align='center',fig.width=16, fig.height=8,dpi=220}

gd_pts <- st_read("../blcn/Pipelines_SHP/Pipelines_GCS_NAD83.shp")%>%clean_names()
#st_crs(gd_pts)

gd_clean<-gd_pts %>%
  filter(substance1 %in% c("Natural Gas","Crude Oil","Hydrocarbon Diluents"))

larp_pop<- st_read("../blcn/LARP/LA_RAC_Recommended_Population_Centers.shp")%>%
  st_transform(crs=st_crs(gd_pts))
larp_bd <-st_read("../blcn/LARP/LUF Integrated Regional Plan Boundaries.shp")%>%
  st_transform(crs=st_crs(gd_pts))%>% st_make_valid()

l_ath_bd <-  st_intersection(gd_clean,larp_bd %>% 
  filter(LUF_CODE=="02"))

ggplot(l_ath_bd)+
  geom_sf(aes(colour=substance1),fill="white") +
  geom_sf(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,15)))+
  geom_sf_text(data = larp_pop%>%filter(OBJECTID %in% c(3,6,7,8,15)), aes(label = NAME))+
  scale_colour_brewer("",palette = "Set1")+
  scale_fill_manual("")+
 theme_void()+
  theme(
    legend.position = "right",
    legend.margin=margin(c(.05,0,.05,0),unit="cm"),
    legend.text = element_text(colour="black", size = 12),
    plot.caption = element_text(size = 10, face = "italic",hjust=0),
    plot.title = element_text(size=16,face = "bold"),
    plot.subtitle = element_text(size = 10),
    #panel.grid.minor = element_blank(),
    text = element_text(size = 20,face = "bold"),
    axis.text.y = element_text(size = 12,face = "bold", colour="black"),
    #axis.text.x = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black", angle = 90,hjust=0.5,vjust=0.5),
    strip.text.x = element_text(size = 8, colour = "black", angle = 0),
    axis.title.y = element_text(size = 14,face = "bold", colour="black"))+
  labs(x="",y="",
      title="Pipeline Infrastructure Within BLCN Territory Over Time",
       #subtitle="Excluding Electricity,by NAICS 4-Digit Code",
       caption="Data via Government of Alberta, map by @andrew_leach")
    

```



```


